---
title: "ICI CKD – Long-term Kidney Outcomes Analysis"
author: "Tianqi Ouyang"
date: "2026-03-19"
format:
  html:
    theme: lumen
    toc: true
    toc-float: true
    toc-depth: 4
    code-fold: true
  pdf:
    toc: true
    toc-depth: 4
execute:
  warning: false
  message: false
---

# 1 Variable Name Dictionary

**Gold standard = master Excel column name** (left column).
`In Code` = what this variable is called in the working code.
Naming rules:

- `ICI_Dose_Date` → `Index_date` (future datasets provide this directly)
- `EMPI` kept uppercase (required by RPDR functions)
- New derived variables: descriptive snake_case
- All other master Excel variable names kept as-is

## 1.1 Identifiers & Demographics

| Master Excel | In Code | Note |
|---|---|---|
| `EMPI` | `EMPI` | Uppercase for RPDR compatibility |
| `Age` | `Age` | |
| `Gender_Legal_Sex` | `Gender_Legal_Sex` | |
| `male` | `male` | 0/1 |
| `Race` | `Race` | |
| `Ethnic_Group` | `Ethnic_Group` | |

## 1.2 Key Dates

| Master Excel | In Code | Note |
|---|---|---|
| `ICI_Dose_Date` | `Index_date` | Renamed; future datasets provide `Index_date` directly |
| `last_cre_date` | `last_cre_date` | Date of last creatinine lab (used as follow-up date) |
| — | `last_cre_days` | Derived: numeric days from Index_date to last_cre_date |

## 1.3 Comorbidities (from master)

| In Code | Description |
|---|---|
| `dm` | Diabetes mellitus |
| `htn` | Hypertension |
| `cad` | Coronary artery disease |
| `esrd_kt` | ESRD / kidney transplant (pre-ICI) |

## 1.4 Medications (from master)

| In Code | Description |
|---|---|
| `ace_arb` | ACE inhibitor / ARB |
| `diu` | Diuretics |
| `ppi` | Proton pump inhibitor |
| `steroids` | Steroids |
| `smoking` | Smoking status |

## 1.5 Cancer Drugs (from master)

| In Code | Description |
|---|---|
| `Bevacizumab` | |
| `Cisplatin` | |
| `Carboplatin` | |
| `Pemetrexed` | |
| `Gemcitabine` | |
| `Cetuximab` | |
| `Trastuzumab` | |
| `VEGF_TKI` | VEGF tyrosine kinase inhibitor |

## 1.6 Pre-baseline Lab Values (from master)

| In Code | Description |
|---|---|
| `pre_CRE_180days` | Creatinine within 180 days pre-ICI |
| `pre_HGB_180days` | Hemoglobin within 180 days pre-ICI |
| `pre_ALB_180days` | Albumin within 180 days pre-ICI |

## 1.7 RPDR-derived Baseline Lab (within 90 days pre-Index_date)

| In Code | RPDR Group_Id | Description |
|---|---|---|
| `pre_median_CRE_90days` | `CRE` | Median creatinine ≤ 90 days before Index_date |

## 1.8 Baseline eGFR & CKD Stage

| In Code | Description |
|---|---|
| `eGFR_CRE_baseline` | CKD-EPI 2021 creatinine eGFR from `pre_median_CRE_90days` |
| `ckd_stage_baseline` | CKD stage 1–5 from baseline eGFR |
| `ckd_stage_baseline_3` | 3-level: "stage 1", "stage 2", "stage 3 or more" |
| `ckd_stage` | From `eGFR_cre` (master column) |

## 1.9 Outcome Variables (from master + derived)

| In Code | Description |
|---|---|
| `ckd_incidence` | CKD incidence (binary, from master) |
| `ckd_incidence_date` | Date of CKD incidence |
| `ckd_progression` | CKD progression (binary, from master) |
| `ckd_progression_date` | Date of CKD progression |
| `esrd_kt_after_ici` | ESRD/KT after ICI (binary, from master) |
| `esrd_kt_after_ici_date` | Date of ESRD/KT after ICI |
| `eskd_defination_1` | ESKD definition 1 (binary, from master) |
| `eskd_defination_1_date` | Date of ESKD definition 1 |
| `eskd_composite` | Derived: esrd_kt_after_ici OR eskd_defination_1 |
| `eskd_composite_date` | pmin of the two ESKD dates |
| `ckd_composite` | Derived: eskd_composite OR ckd_incidence OR ckd_progression |
| `time_to_ckd_composite` | pmin of all three time-to-event values |

## 1.10 Time-to-Event Variables

| In Code | Description |
|---|---|
| `time_to_eskd_composite` | Days from Index_date to ESKD composite |
| `time_to_ckd_incidence` | Days from Index_date to CKD incidence |
| `time_to_ckd_progression` | Days from Index_date to CKD progression |
| `time_to_ckd_composite` | pmin of the above three |

## 1.11 Survival Flags (1–10 years)

| Pattern | Description |
|---|---|
| `survival_Xyear` | 1 if last_cre_days ≥ 365 × X |

## 1.12 Yearly Outcome Flags (1–10 years)

| Pattern | Description |
|---|---|
| `ckd_incidence_Xyear` | 1 if time_to_ckd_incidence < 365 × X |
| `ckd_progression_Xyear` | 1 if time_to_ckd_progression < 365 × X |
| `eskd_composite_Xyear` | 1 if time_to_eskd_composite < 365 × X |

## 1.13 Yearly eGFR & CKD Stage (from RPDR lab, post-Index_date)

| Pattern | Description |
|---|---|
| `pre_median_CRE_Xyear` | Median CRE in 90-day window around year X post-Index |
| `eGFR_CRE_Xyear` | CKD-EPI eGFR from that median CRE |
| `ckd_stage_Xyear` | CKD stage 1–5 from that eGFR |
| `lab_num_Xyear` | Number of CRE labs in that yearly window |

## 1.14 Mortality (from master)

| In Code | Description |
|---|---|
| `death_1year` … `death_10year` | Binary death flags at each year mark |

---

# 2 Setup & Libraries

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#| label: libraries
rm(list = ls())
library(readxl)
library(dplyr)
library(tidyverse)
library(tableone)
library(writexl)
library(gtsummary)
library(survival)
library(ggplot2)
library(ggpubr)
library(splines)
library(lattice)
library(mice)
library(DescTools)
library(survminer)
library(finalfit)
library(tidycmprsk)
library(ggsurvfit)
library(car)
library(cobalt)
library(pammtools)
library(adjustedCurves)
library(rms)
library(riskRegression)
library(prodlim)
library(bodycompref)
library(table1)
library(readr)
```

---

# 3 Functions

Each function is defined in its own labeled cell.

## 3.1 pre_lab_median

Gets the **median** lab value within a time window **before** `index_date`.
Returns one row per patient with the median result.

```{r}
#| label: fn-pre-lab-median
pre_lab_median <- function(data, month, time, index_date, Group_Id) {
  index_date <- enquo(index_date)
  if (month) {
    time_value <- time * 30
  } else {
    time_value <- time
  }

  median_name <- paste0("pre_median_", Group_Id, "_", time_value, "days")

  result <- data %>%
    filter(Group_Id == !!Group_Id, grepl("^\\d+\\.\\d+$", Result)) %>%
    filter(!is.na(Result)) %>%
    mutate(Result = as.numeric(Result)) %>%
    filter(Seq_Date_Time >= (!!index_date) - time_value,
           Seq_Date_Time <= !!index_date) %>%
    group_by(EMPI) %>%
    summarise(
      !!median_name := median(Result, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    select(EMPI, !!median_name)

  return(result)
}
```

## 3.2 pre_lab_median_year

Gets the **median** CRE in a 90-day window around each yearly mark post-index.
Also finds the **closest** lab value up to that yearly mark as a fallback.
Falls back to the closest value if no labs exist in the 90-day window.

```{r}
#| label: fn-pre-lab-median-year
pre_lab_median_year <- function(data, year, days, index_date, Group_Id) {
  index_date <- enquo(index_date)

  last_offset  <- year * 365
  first_offset <- last_offset - days

  median_name      <- paste0("pre_median_", Group_Id, "_", year, "year")
  lab_num_name     <- paste0("lab_num_", year, "year")
  closest_name     <- paste0("closest_", Group_Id, "_", year, "year")
  closest_date_name <- paste0("closest_", Group_Id, "_", year, "year_date")

  median <- data %>%
    filter(Group_Id == !!Group_Id, grepl("^\\d+\\.\\d+$", Result)) %>%
    filter(!is.na(Result)) %>%
    mutate(Result = as.numeric(Result)) %>%
    filter(Seq_Date_Time >= (!!index_date + first_offset),
           Seq_Date_Time <= !!index_date + last_offset) %>%
    group_by(EMPI) %>%
    summarise(
      !!median_name := median(Result, na.rm = TRUE),
      !!lab_num_name := n(),
      .groups = "drop"
    )

  closest <- data %>%
    filter(Group_Id == !!Group_Id, grepl("^\\d+\\.\\d+$", Result)) %>%
    filter(!is.na(Result)) %>%
    mutate(Result = as.numeric(Result)) %>%
    filter(Seq_Date_Time <= !!index_date + last_offset) %>%
    arrange(EMPI, desc(Seq_Date_Time)) %>%
    distinct(EMPI, .keep_all = TRUE) %>%
    rename(!!closest_name := Result,
           !!closest_date_name := Seq_Date_Time) %>%
    select(EMPI, !!index_date, !!closest_name, !!closest_date_name)

  result <- median %>%
    merge(closest, by = "EMPI", all = TRUE) %>%
    mutate(
      !!median_name := ifelse(is.na(!!sym(median_name)),
                              !!sym(closest_name),
                              !!sym(median_name))
    ) %>%
    select(EMPI, !!median_name, !!lab_num_name)

  return(result)
}
```

## 3.3 left_join_all

Sequentially left-joins a list of data frames onto a master data frame.

```{r}
#| label: fn-left-join-all
left_join_all <- function(master, df_list, by) {
  for (df in df_list) {
    master <- merge(master, df, by = by, all.x = TRUE)
  }
  return(master)
}
```

## 3.4 get_ckd_stage

Classifies CKD stages 1–5 based on eGFR values.

```{r}
#| label: fn-get-ckd-stage
get_ckd_stage <- function(egfr) {
  if (is.na(egfr)) {
    return(NA)
  } else if (egfr >= 90) {
    return(1)
  } else if (egfr >= 60) {
    return(2)
  } else if (egfr >= 30) {
    return(3)
  } else if (egfr >= 15) {
    return(4)
  } else {
    return(5)
  }
}
```

## 3.5 ckd_epi_gfr (CKD-EPI 2021 creatinine-only)

Race-free CKD-EPI 2021 creatinine-based eGFR.

```{r}
#| label: fn-ckd-epi-gfr
ckd_epi_gfr <- function(creat, male, age) {
  k <- ifelse(male == 1, 0.9, 0.7)
  a <- ifelse(male == 1, -0.302, -0.241)
  m <- ifelse(male == 1, 1, 1.012)
  gfr <- 142 * min(c(creat / k, 1))^a * max(c(creat / k, 1))^-1.2 * 0.9938^age * m
  return(gfr)
}

gfr_calc <- Vectorize(ckd_epi_gfr)
```

## 3.6 create_outcome_table

Generates a `gtsummary` Table 1 for a given year suffix, filtering to patients
who survived to that landmark.

```{r}
#| label: fn-create-outcome-table
create_outcome_table <- function(data, time_suffix) {

  year_num <- as.numeric(gsub("[^0-9]", "", time_suffix))
  stage_history_cols <- paste0("ckd_stage_", 1:year_num, "year")

  survival_col     <- paste0("survival_", time_suffix)
  ckd_inc_col      <- paste0("ckd_incidence_", time_suffix)
  ckd_prog_col     <- paste0("ckd_progression_", time_suffix)
  eskd_comp_col    <- paste0("eskd_composite_", time_suffix)
  ckd_comp_col     <- paste0("ckd_composite_", time_suffix)

  data %>%
    filter(.data[[survival_col]] == 1) %>%
    select(
      ## Demographics
      Age, male, Race, Ethnic_Group,
      ## Outcomes (dynamic)
      all_of(ckd_inc_col), all_of(ckd_prog_col),
      all_of(eskd_comp_col), all_of(ckd_comp_col),
      ## Longitudinal CKD Stages (cumulative up to year X)
      all_of(stage_history_cols),
      ckd_stage_baseline,
      aki,
      ## Comorbidities
      dm, htn, cad, cirrhosis,
      ## Medications
      ace_arb, diu, ppi, steroids, smoking, statins,
      ICI_Class, nephrotoxic_chemo,
      ## Labs / Kidney function
      pre_CRE_180days, pre_HGB_180days, pre_ALB_180days,
      eGFR_CRE_baseline
    ) %>%
    tbl_summary(missing_text = "Missing")
}
```

## 3.7 pre_lab (closest value — Cystatin C style)

Finds the **closest lab value on or before** `index_date` within the time window.
Returns one row per patient (most recent qualifying lab result).

> Uses `Seq_Date_Time` for filtering and sorting (bug-fixed version from Cystatin C pipeline).

```{r}
#| label: fn-pre-lab
pre_lab <- function(data, month, time, index_date, Group_Id) {
  index_date <- enquo(index_date)
  if (month) {
    time_value <- time * 30
  } else {
    time_value <- time
  }

  lab_name      <- paste0("pre_", Group_Id, "_", time_value, "days")
  lab_date_name <- paste0("pre_", Group_Id, "_", time_value, "days_date")

  result <- data %>%
    filter(Group_Id == !!Group_Id, grepl("^\\d+\\.\\d+$", Result)) %>%
    filter(!is.na(Result)) %>%
    mutate(Result = as.numeric(Result)) %>%
    filter(Seq_Date_Time >= (!!index_date) - time_value,
           Seq_Date_Time <= !!index_date) %>%
    arrange(EMPI, desc(Seq_Date_Time)) %>%
    distinct(EMPI, .keep_all = TRUE) %>%
    rename(!!lab_name := Result,
           !!lab_date_name := Seq_Date_Time) %>%
    select(EMPI, !!lab_name, !!lab_date_name)

  return(result)
}
```

## 3.8 dia_list_name (RPDR diagnosis lookup)

Identifies patients with a given diagnosis (and optionally medication) on or before
`index_date`. Supports excluding specific diagnosis name strings.

```{r}
#| label: fn-dia-list-name
dia_list_name <- function(data, listed_names, index_date, excluded_names = NULL, med_data = NULL, med_names = NULL) {
  if (!is.null(excluded_names)) {
    excluded_pattern <- paste(
      gsub("([][()])", "\\\\\\1", excluded_names),
      collapse = "|"
    )
  }

  if (is.null(excluded_names)) {
    dia_result <- data %>%
      filter(Date <= {{index_date}}) %>%
      filter(grepl(paste(listed_names, collapse = "|"), Diagnosis_Name, ignore.case = TRUE)) %>%
      distinct(EMPI)
  } else {
    dia_result <- data %>%
      filter(Date <= {{index_date}}) %>%
      filter(
        grepl(paste(listed_names, collapse = "|"), Diagnosis_Name, ignore.case = TRUE),
        !grepl(excluded_pattern, Diagnosis_Name, ignore.case = TRUE)
      ) %>%
      distinct(EMPI)
  }

  if (!is.null(med_data)) {
    med_result <- med_data %>%
      filter(Medication_Date <= {{index_date}}) %>%
      filter(grepl(paste(med_names, collapse = "|"), Medication, ignore.case = TRUE)) %>%
      distinct(EMPI)
  } else {
    med_result <- NULL
  }

  final_result <- rbind(dia_result, med_result) %>%
    distinct(EMPI)

  return(final_result)
}
```

## 3.9 cmp_variable (competing risk)

Builds competing-risk time and status for a single outcome.
Status coding: 0 = censored, 1 = event occurred, 2 = death (competing event).

```{r}
#| label: fn-cmp-variable
cmp_variable <- function(data, event_time, event_status, name) {
  event_time   <- enquo(event_time)
  event_status <- enquo(event_status)

  cmp_time_name   <- paste0("cmp_time_", name)
  cmp_status_name <- paste0("cmp_status_", name)

  data %>%
    mutate(
      !!cmp_time_name   := ifelse(time_to_death <= !!event_time, time_to_death, !!event_time),
      !!cmp_status_name := ifelse(time_to_death <= !!event_time, time_to_death_status, !!event_status)
    )
}
```

## 3.10 make_ckd_outcome_date (sustained eGFR-based outcomes)

Identifies sustained (≥ days_threshold) episodes where a condition is met.
Returns one row per patient with binary flag + first qualifying date.

```{r}
#| label: fn-make-ckd-outcome
make_ckd_outcome_date <- function(data, id_var, date_var, condition_expr,
                                   days_threshold = 90, outcome_name = "default") {
  id_var      <- enquo(id_var)
  date_var    <- enquo(date_var)
  cond_q      <- enquo(condition_expr)
  outcome_sym <- ensym(outcome_name)
  outcome_date_name <- paste0(outcome_name, "_date")

  working <- data %>%
    arrange(!!id_var, !!date_var) %>%
    group_by(!!id_var) %>%
    mutate(
      .cond   = if_else(!!cond_q, 1L, 0L),
      .change = .cond != lag(.cond, default = first(.cond)),
      .run    = cumsum(.change)
    )

  episodes <- working %>%
    group_by(!!id_var, .run) %>%
    summarise(
      .code = first(.cond),
      .start = first(!!date_var),
      .last  = last(!!date_var),
      .groups = "drop"
    ) %>%
    group_by(!!id_var) %>%
    mutate(
      .end  = lead(.start),
      .end  = if_else(is.na(.end), .last, .end),
      .days = as.integer(.end - .start)
    ) %>%
    ungroup()

  episodes %>%
    filter(!is.na(.days)) %>%
    group_by(!!id_var) %>%
    summarise(
      has_event  = any(.code == 1L & .days >= days_threshold, na.rm = TRUE),
      event_date = if (any(.code == 1L & .days >= days_threshold, na.rm = TRUE)) {
        min(.start[.code == 1L & .days >= days_threshold], na.rm = TRUE)
      } else {
        as.Date(NA)
      },
      .groups = "drop"
    ) %>%
    mutate(!!outcome_sym := as.integer(has_event)) %>%
    rename(!!outcome_date_name := event_date) %>%
    select(-has_event)
}
```

## 3.12 med_list (RPDR medication lookup)

```{r}
#| label: fn-med-list
med_list <- function(names, index_date, excluded_names = NULL, data) {
  if (!is.null(excluded_names)) {
    result <- data %>%
      filter(Medication_Date <= {{index_date}}) %>%
      filter(
        grepl(paste(names, collapse = "|"), Medication, ignore.case = TRUE),
        !grepl(paste(excluded_names, collapse = "|"), Medication, ignore.case = TRUE)
      ) %>%
      distinct(EMPI)
  } else {
    result <- data %>%
      filter(Medication_Date <= {{index_date}}) %>%
      filter(grepl(paste(names, collapse = "|"), Medication, ignore.case = TRUE)) %>%
      distinct(EMPI)
  }
  return(result)
}
```

## 3.13 dia_list_icd (RPDR diagnosis lookup by ICD code)

```{r}
#| label: fn-dia-list-icd
dia_list_icd <- function(data, index_date, icd_code, excluded_names = NULL,
                         med_data = NULL, med_names = NULL) {
  if (is.null(excluded_names)) {
    dia_result <- data %>%
      filter(Date <= {{index_date}}) %>%
      filter(grepl(paste(icd_code, collapse = "|"), Code, ignore.case = TRUE)) %>%
      distinct(EMPI)
  } else {
    dia_result <- data %>%
      filter(Date <= {{index_date}}) %>%
      filter(
        grepl(paste(icd_code, collapse = "|"), Code, ignore.case = TRUE),
        !grepl(paste(excluded_names, collapse = "|"), Diagnosis_Name, ignore.case = TRUE)
      ) %>%
      distinct(EMPI)
  }

  if (!is.null(med_data)) {
    med_result <- med_data %>%
      filter(Medication_Date <= {{index_date}}) %>%
      filter(grepl(paste(med_names, collapse = "|"), Medication, ignore.case = TRUE)) %>%
      distinct(EMPI)
  } else {
    med_result <- NULL
  }

  final_result <- rbind(dia_result, med_result) %>% distinct(EMPI)
  return(final_result)
}
```

## 3.14 dia_list_icd_after_index (post-index diagnosis with first event date)

```{r}
#| label: fn-dia-list-icd-after
dia_list_icd_after_index <- function(data, index_date, icd_code,
                                     excluded_names = NULL) {
  if (is.null(excluded_names)) {
    result <- data %>%
      filter(Date > {{index_date}}) %>%
      filter(grepl(paste(icd_code, collapse = "|"), Code, ignore.case = TRUE))
  } else {
    result <- data %>%
      filter(Date > {{index_date}}) %>%
      filter(
        grepl(paste(icd_code, collapse = "|"), Code, ignore.case = TRUE),
        !grepl(paste(excluded_names, collapse = "|"), Diagnosis_Name, ignore.case = TRUE)
      )
  }

  result %>%
    arrange(EMPI, Date) %>%
    distinct(EMPI, .keep_all = TRUE) %>%
    select(EMPI, first_event_date = Date)
}
```

---

# 4 Data Loading

## 4.1 Load master file

> **Future datasets** will provide `EMPI` and `Index_date` columns directly.
> Update the path and column rename below accordingly.

```{r}
#| label: load-master
# ── Update this path to your master Excel file ──
# Drop pre-computed columns — we derive them from RPDR Dia/Med/Lab
# Drop pre-computed columns — we derive them from RPDR
drop_cols <- c("eGFR_cre", "ckd_stage", "last_date", "last_days",
               "dm", "htn", "cad", "ace_arb", "diu", "ppi", "steroids",
               "smoking", "esrd_kt", "esrd_kt_after_ici", "esrd_kt_after_ici_date",
               "ckd_incidence", "ckd_incidence_date", "ckd_progression", "ckd_progression_date",
               "eskd_defination_1", "eskd_defination_1_date",
               "ckd_composite", "time_to_ckd_incidence", "time_to_ckd_progression",
               "time_to_ckd_composite",
               paste0("ckd_incidence_", 1:10, "year"),
               paste0("ckd_progression_", 1:10, "year"),
               paste0("ckd_composite_", 1:10, "year"),
               paste0("death_", 1:10, "year"))
final_master <- read_excel(
  "/Users/to909/Desktop/Meg/Kavita ICI CKD/final_master_1217.xlsx"
) %>%
  select(-any_of(drop_cols))

# ── Rename ICI_Dose_Date → Index_date for future compatibility ──
# If your dataset already has Index_date, remove this rename.
final_master <- final_master %>%
  rename(Index_date = ICI_Dose_Date)
```

## 4.2 Load RPDR files (Lab, Dia, Med)

```{r}
#| label: load-rpdr
rpdr_path <- "/Users/to909/Partners HealthCare Dropbox/Tianqi Ouyang/CKD after ICI/RPDR/"

# Lab
lab <- read.table(
  paste0(rpdr_path, "mes98_093025104659807332_Lab.txt"),
  sep = "|", fill = TRUE, quote = "", header = TRUE
) %>%
  select(EMPI, EPIC_PMRN, Seq_Date_Time, Result, Group_Id, Test_Description)
lab$Seq_Date_Time <- as.Date(lab$Seq_Date_Time, "%m/%d/%Y")

# Diagnoses
dia <- read.table(
  paste0(rpdr_path, "mes98_093025104659807332_Dia.txt"),
  sep = "|", fill = TRUE, quote = "", header = TRUE
) %>%
  select(EMPI, EPIC_PMRN, Code, Date, Diagnosis_Name)
dia$Date <- as.Date(dia$Date, "%m/%d/%Y")

# Medications
med <- read.table(
  paste0(rpdr_path, "mes98_093025104659807332_Med.txt"),
  sep = "|", fill = TRUE, quote = "", header = TRUE
) %>%
  select(EMPI, Medication_Date, Medication)
med$Medication_Date <- as.Date(med$Medication_Date, "%m/%d/%Y")
```

## 4.3 Coerce dates & create master_lab / master_dia / master_med

```{r}
#| label: load-dates
final_master$Index_date    <- as.Date(final_master$Index_date)
final_master$last_cre_date <- as.Date(final_master$last_cre_date)

master_list <- final_master %>% select(EMPI, Index_date)
master_lab  <- lab %>% merge(master_list, by = "EMPI")
master_dia  <- dia %>% merge(master_list, by = "EMPI")
master_med  <- med %>% merge(master_list, by = "EMPI")
```

---

# 5 Comorbidities & Medications (from RPDR Dia + Med)

## 5.1 Medication name lists

```{r}
#| label: med-names
ppi_names <- c("omeprazole", "Prilosec", "Yosprala", "lansoprazole", "Prevacid",
               "dexlansoprazole", "Dexilent", "rabeprazole", "Aciphex",
               "pantoprazole", "Protonix", "esomeprazole", "Nexium", "Vimovo",
               "Zegerid", "dexilant")

ace_arb_names <- c("benazepril", "lotensin", "captopril", "capoten", "enalapril",
                   "vasotech", "epaned", "lexxel", "fosinopril", "monopril",
                   "lisinopril", "prinvil", "zestril", "qbrelis", "moexipril",
                   "univasc", "perindopril", "aceon", "quinapril", "accupril",
                   "ramipril", "altace", "trandolapril", "mavik", "azilsartan",
                   "edarbi", "candesartan", "atacand", "eprosartan", "teveten",
                   "irbesartan", "avapro", "telmisartan", "micardis", "valsartan",
                   "diovan", "losartan", "cozaar", "olmesartan", "benicar",
                   "lotrel", "hyzaar")

diuretics_names <- c("bumetanide", "bumex", "ethacrynic acid", "edecrin",
                     "furosemide", "Lasix", "torsemide", "demadex",
                     "hydrochlorothiazide", "hctz", "chlorathiazide",
                     "chlorthalidone", "diuril", "indapamide", "metolazone",
                     "aldactone", "amiloride", "dyazide", "eplerenone",
                     "maxzide", "spironolactone", "triamterene")

steroids_names <- c("prednisone", "methylpred", "Deltasone", "Orasone",
                    "budesonide", "Entocort", "methylprednisolone",
                    "prednisolone", "Millipred", "dexamethasone", "Ozurdex",
                    "Maxidex", "DexPak")
no_steroid_names <- c("Inhaler", "Nebulization", "Ointment", "Ophthalmic",
                      "Symbicort")

statins_names <- c("Atorvastatin", "Lipitor", "Fluvastatin", "Lescol",
                   "Lovastatin", "Mevacor", "Pitavastatin", "Livalo",
                   "Pravastatin", "Pravachol", "Rosuvastatin", "Crestor",
                   "Simvastatin", "Zocor")
```

## 5.2 Diagnosis code lists

```{r}
#| label: dia-codes
esrd_kt_codes <- c(
  "585.6", "N18.6", "V45.1", "Z94.0", "Z99.2",
  "996.81", "T86.1", "T86.10", "T86.11", "T86.12", "T86.13", "T86.19",
  "N18.0", "V56.1", "Z49.0", "Z49.1", "Z49.2"
)

diabetes_codes <- c(
  "250.00", "250.01", "250.02", "250.03", "250.10", "250.11", "250.12",
  "250.13", "250.40", "250.41", "250.42", "250.43",
  "E10.9", "E11.9", "E10.65", "E11.65", "E10.21", "E11.21",
  "E10.22", "E11.22", "E10.29", "E11.29", "E10.40", "E11.40",
  "E10.49", "E11.49", "E13.9"
)

htn_codes <- c("401.1", "401.0", "401.9", "I10", "I15.1", "I15.9",
               "I15.8", "I15.0", "I15.2", "405.01", "405.09", "405.91",
               "405.99")

no_htn_list <- c("Pulmonary Embolism", "Borderline hypertension",
                 "Labile hypertension", "Chronic venous hypertension")

htn_med_list <- c("chlorthalidone", "chlorothiazide", "hydrochlorothiazide",
                  "indapamide", "metolazone", "spironolactone", "triamterene",
                  "amlodipine", "diltiazem", "felodipine", "nifedipine",
                  "verapamil", "benazepril", "captopril", "enalapril",
                  "fosinopril", "lisinopril", "ramipril", "losartan",
                  "valsartan", "irbesartan", "olmesartan", "candesartan",
                  "clonidine", "hydralazine", "minoxidil")

cad_list <- c("Coronary artery disease", "Coronary atherosclerosis",
              "Atherosclerotic heart disease", "Angina pectoris",
              "Chronic ischemic heart disease", "Unstable angina",
              "Ischemic heart disease", "Ischemic cardiomyopathy",
              "Coronary occlusion without myocardial infarction")

smoking_list <- c("smoking", "tobacco", "cigarette", "cigarettes",
                  "nicotine", "smoke")
```

## 5.3 Derive comorbidities & medications from RPDR

```{r}
#| label: derive-comorbidities
# Medications
ppi      <- med_list(data = master_med, names = ppi_names, index_date = Index_date)
ace_arb  <- med_list(data = master_med, names = ace_arb_names, index_date = Index_date)
diu      <- med_list(data = master_med, names = diuretics_names, index_date = Index_date)
steroids <- med_list(data = master_med, names = steroids_names,
                     index_date = Index_date, excluded_names = no_steroid_names)
statins  <- med_list(data = master_med, names = statins_names, index_date = Index_date)

# Diagnoses
esrd_kt <- dia_list_icd(master_dia, icd_code = esrd_kt_codes, index_date = Index_date)
esrd_kt_after_ici <- dia_list_icd_after_index(master_dia, icd_code = esrd_kt_codes,
                                               index_date = Index_date) %>%
  rename(esrd_kt_after_ici_date = first_event_date)
dm  <- dia_list_icd(master_dia, icd_code = diabetes_codes, index_date = Index_date)
htn <- dia_list_icd(master_dia, icd_code = htn_codes, excluded_names = no_htn_list,
                    med_data = master_med, med_names = htn_med_list,
                    index_date = Index_date)
cad <- dia_list_name(master_dia, listed_names = cad_list, index_date = Index_date)
cirrhosis <- master_dia %>%
  filter(Date <= Index_date) %>%
  filter(grepl("cirrhosis", Diagnosis_Name, ignore.case = TRUE)) %>%
  distinct(EMPI)
smoking <- master_dia %>%
  filter(grepl(paste(smoking_list, collapse = "|"), Diagnosis_Name, ignore.case = TRUE)) %>%
  distinct(EMPI)

# Join all comorbidities/medications to final_master
final_master <- final_master %>%
  mutate(
    diu      = ifelse(EMPI %in% diu$EMPI, 1, 0),
    ace_arb  = ifelse(EMPI %in% ace_arb$EMPI, 1, 0),
    esrd_kt  = ifelse(EMPI %in% esrd_kt$EMPI, 1, 0),
    esrd_kt_after_ici = ifelse(EMPI %in% esrd_kt_after_ici$EMPI, 1, 0),
    dm       = ifelse(EMPI %in% dm$EMPI, 1, 0),
    htn      = ifelse(EMPI %in% htn$EMPI, 1, 0),
    ppi      = ifelse(EMPI %in% ppi$EMPI, 1, 0),
    steroids = ifelse(EMPI %in% steroids$EMPI, 1, 0),
    smoking  = ifelse(EMPI %in% smoking$EMPI, 1, 0),
    cad      = ifelse(EMPI %in% cad$EMPI, 1, 0),
    cirrhosis = ifelse(EMPI %in% cirrhosis$EMPI, 1, 0),
    statins  = ifelse(EMPI %in% statins$EMPI, 1, 0)
  )

# Merge esrd_kt_after_ici date
final_master <- final_master %>%
  merge(esrd_kt_after_ici, by = "EMPI", all.x = TRUE)
```

## 5.4 ICI_Class & nephrotoxic_chemo

```{r}
#| label: derive-ici-class
final_master <- final_master %>%
  mutate(
    ICI_Class = case_when(
      ICI_Name %in% c("pembrolizumab", "nivolumab", "cemiplimab",
                       "Nivolumab", "Dostarlimab", "Cemiplimab") ~ "PD1",
      ICI_Name %in% c("atezolizumab", "avelumab", "durvalumab") ~ "PDL1",
      ICI_Name == "ipilimumab" ~ "CTLA4",
      ICI_Name == "ipilimumab AND nivolumab" ~ "Combination",
      TRUE ~ NA_character_
    ),
    nephrotoxic_chemo = if_else(
      if_any(
        any_of(c("Bevacizumab", "Cisplatin", "Carboplatin", "Pemetrexed",
                  "Gemcitabine", "Cetuximab", "Trastuzumab")),
        ~ . == 1
      ),
      1L, 0L
    )
  )
```

---

# 6 Lab-based Variable Creation

## 6.1 Baseline creatinine from RPDR (median within 90 days pre-index)

```{r}
#| label: clean-baseline-cre
pre_lab_baseline <- pre_lab_median(
  master_lab, month = FALSE, time = 90,
  index_date = Index_date, Group_Id = "CRE"
)

final_master <- final_master %>%
  merge(pre_lab_baseline, by = "EMPI", all.x = TRUE)
```

## 6.2 Baseline eGFR & CKD stage

```{r}
#| label: clean-baseline-egfr
final_master$eGFR_CRE_baseline <- mapply(
  gfr_calc,
  final_master$pre_median_CRE_90days,
  final_master$male,
  final_master$Age
)

final_master <- final_master %>%
  mutate(
    ckd_stage_baseline = case_when(
      eGFR_CRE_baseline >= 90                           ~ 1,
      eGFR_CRE_baseline >= 60 & eGFR_CRE_baseline < 90 ~ 2,
      eGFR_CRE_baseline >= 30 & eGFR_CRE_baseline < 60 ~ 3,
      eGFR_CRE_baseline >= 15 & eGFR_CRE_baseline < 30 ~ 4,
      eGFR_CRE_baseline < 15                            ~ 5
    ),
    ckd_stage_baseline_3 = case_when(
      ckd_stage_baseline == 1  ~ "stage 1",
      ckd_stage_baseline == 2  ~ "stage 2",
      ckd_stage_baseline >= 3  ~ "stage 3 or more"
    )
  )
```

## 6.3 CKD outcomes from RPDR Lab (sustained eGFR-based)

```{r}
#| label: clean-ckd-outcomes
# Build post-baseline eGFR series for each patient
outcome_labs <- master_lab %>%
  select(EMPI, Seq_Date_Time, Result, Group_Id, Index_date) %>%
  merge(final_master %>% select(EMPI, pre_median_CRE_90days, eGFR_CRE_baseline,
                                 male, Age),
        by = "EMPI") %>%
  filter(Group_Id == "CRE", grepl("^\\d+\\.\\d+$", Result)) %>%
  filter(!is.na(Result)) %>%
  mutate(Result = as.numeric(Result)) %>%
  filter(Seq_Date_Time > Index_date) %>%
  mutate(eGFR_CRE = mapply(gfr_calc, Result, male, Age)) %>%
  select(EMPI, eGFR_CRE_baseline, eGFR_CRE, eGFR_CRE_date = Seq_Date_Time) %>%
  mutate(
    pct_drop = round((eGFR_CRE_baseline - eGFR_CRE) / eGFR_CRE_baseline, 3)
  ) %>%
  arrange(EMPI, eGFR_CRE_date)

# CKD incidence: >30% drop AND eGFR <60, from baseline >=60, sustained >=90 days
ckd_incidence <- make_ckd_outcome_date(
  data = outcome_labs, id_var = EMPI, date_var = eGFR_CRE_date,
  condition_expr = pct_drop > 0.30 & eGFR_CRE < 60 & eGFR_CRE_baseline >= 60,
  days_threshold = 90, outcome_name = "ckd_incidence"
)

# ESKD definition 1: eGFR <10 sustained >=90 days
eskd_defination_1 <- make_ckd_outcome_date(
  data = outcome_labs, id_var = EMPI, date_var = eGFR_CRE_date,
  condition_expr = eGFR_CRE < 10,
  days_threshold = 90, outcome_name = "eskd_defination_1"
)

# CKD progression: eGFR <60 from baseline <60, sustained >=90 days
ckd_progression <- make_ckd_outcome_date(
  data = outcome_labs, id_var = EMPI, date_var = eGFR_CRE_date,
  condition_expr = eGFR_CRE < 60 & eGFR_CRE_baseline < 60,
  days_threshold = 90, outcome_name = "ckd_progression"
)

# Merge outcomes
final_master <- final_master %>%
  merge(ckd_incidence, by = "EMPI", all.x = TRUE) %>%
  merge(eskd_defination_1, by = "EMPI", all.x = TRUE) %>%
  merge(ckd_progression, by = "EMPI", all.x = TRUE) %>%
  mutate(
    ckd_incidence      = replace_na(ckd_incidence, 0),
    eskd_defination_1  = replace_na(eskd_defination_1, 0),
    ckd_progression    = replace_na(ckd_progression, 0),
    esrd_kt_after_ici  = replace_na(esrd_kt_after_ici, 0),
    eskd_composite     = ifelse(esrd_kt_after_ici == 1 | eskd_defination_1 == 1, 1, 0),
    eskd_composite_date = pmin(esrd_kt_after_ici_date, eskd_defination_1_date, na.rm = TRUE)
  )

# Coerce outcome dates
final_master <- final_master %>%
  mutate(across(
    any_of(c("esrd_kt_after_ici_date", "eskd_defination_1_date",
             "ckd_incidence_date", "ckd_progression_date",
             "eskd_composite_date", "Index_date", "last_cre_date")),
    as.Date
  ))
```

## 6.4 Time-to-event, composites, survival flags, last_cre_days

```{r}
#| label: clean-time-to-event
final_master <- final_master %>%
  mutate(
    time_to_eskd_composite  = as.numeric(eskd_composite_date  - Index_date),
    time_to_ckd_incidence   = as.numeric(ckd_incidence_date   - Index_date),
    time_to_ckd_progression = as.numeric(ckd_progression_date - Index_date),
    ckd_composite = ifelse(
      eskd_composite == 1 | ckd_incidence == 1 | ckd_progression == 1, 1, 0
    ),
    time_to_ckd_composite = pmin(
      time_to_eskd_composite, time_to_ckd_incidence,
      time_to_ckd_progression, na.rm = TRUE
    ),
    last_cre_days = as.numeric(last_cre_date - Index_date),
    death_days    = as.numeric(Date_Of_Death - Index_date),
    time_to_death = death_days,
    time_to_death_status = ifelse(!is.na(death_days), 2, 0),
    death         = ifelse(!is.na(Date_Of_Death), 1, 0),
    last_date     = if_else(is.na(Date_Of_Death), last_cre_date, Date_Of_Death),
    last_days     = as.numeric(last_date - Index_date)
  ) %>%
  filter(last_cre_days > 0)
```

## 6.5 Death & survival flags (1–10 years)

```{r}
#| label: clean-survival-flags
for (yr in 1:10) {
  # death_Xyear: last_days < 365*X (matches Kavita master builder)
  final_master[[paste0("death_", yr, "year")]] <-
    ifelse(final_master$last_days < 365 * yr, 1, 0)
  # survival_Xyear: last_cre_days >= 365*X (lab follow-up)
  col_name <- paste0("survival_", yr, "year")
  final_master[[col_name]] <- ifelse(final_master$last_cre_days >= 365 * yr, 1, 0)
}
```

## 6.6 Yearly outcome flags (1–10 years)

```{r}
#| label: clean-yearly-outcomes
for (yr in 1:10) {
  final_master[[paste0("ckd_incidence_", yr, "year")]] <-
    ifelse(final_master$time_to_ckd_incidence < 365 * yr, 1, 0)

  final_master[[paste0("ckd_progression_", yr, "year")]] <-
    ifelse(final_master$time_to_ckd_progression < 365 * yr, 1, 0)

  final_master[[paste0("eskd_composite_", yr, "year")]] <-
    ifelse(final_master$time_to_eskd_composite < 365 * yr, 1, 0)

  final_master[[paste0("ckd_composite_", yr, "year")]] <-
    ifelse(final_master$time_to_ckd_composite < 365 * yr, 1, 0)
}

# Replace NA with 0, then convert to factor
final_master <- final_master %>%
  mutate(across(
    .cols = matches("(ckd_incidence|ckd_progression|eskd_composite|ckd_composite).*year$"),
    .fns  = ~ factor(replace_na(., 0))
  ))
```

## 6.7 Yearly CRE medians, eGFR, CKD stage (1–10 years post-index, from RPDR)

```{r}
#| label: clean-yearly-cre
# Compute yearly median CRE (90-day window around each year mark)
yearly_cre_list <- list()
for (yr in 1:10) {
  yearly_cre_list[[yr]] <- pre_lab_median_year(
    master_lab, year = yr, days = 90,
    index_date = Index_date, Group_Id = "CRE"
  )
}

final_master <- left_join_all(final_master, yearly_cre_list, by = "EMPI")

# Compute eGFR and CKD stage for each year
for (yr in 1:10) {
  cre_col   <- paste0("pre_median_CRE_", yr, "year")
  egfr_col  <- paste0("eGFR_CRE_", yr, "year")
  stage_col <- paste0("ckd_stage_", yr, "year")

  final_master[[egfr_col]]  <- mapply(gfr_calc, final_master[[cre_col]],
                                       final_master$male, final_master$Age)
  final_master[[stage_col]] <- sapply(final_master[[egfr_col]], get_ckd_stage)
}
```

## 6.8 AKI (from RPDR Lab)

```{r}
#| label: clean-aki
# AKI: post-baseline CRE > 1.5 × pre-baseline CRE within 90 days (KDIGO grade 2)
aki_data <- master_lab %>%
  merge(pre_lab_baseline, by = "EMPI", all.x = TRUE) %>%
  filter(Group_Id == "CRE", grepl("^\\d+\\.\\d+$", Result)) %>%
  filter(!is.na(Result), !is.na(pre_median_CRE_90days)) %>%
  mutate(Result = as.numeric(Result)) %>%
  filter(Seq_Date_Time > Index_date,
         Seq_Date_Time <= Index_date + 90) %>%
  filter(Result > 1.5 * pre_median_CRE_90days) %>%
  arrange(EMPI, Seq_Date_Time) %>%
  distinct(EMPI, .keep_all = TRUE) %>%
  select(EMPI, aki_date = Seq_Date_Time)

final_master <- final_master %>%
  merge(aki_data, by = "EMPI", all.x = TRUE) %>%
  mutate(
    aki = ifelse(!is.na(aki_date), 1, 0),
    aki_days = as.numeric(aki_date - Index_date),
    aki_before_ckd = ifelse(aki_days <= time_to_ckd_composite, 1, 0)
  )
```

## 6.9 Cohort exclusion filters

```{r}
#| label: clean-exclusion
# Exclude: missing baseline CRE, prior ESRD/KT, baseline eGFR < 10
final_master <- final_master %>%
  filter(!is.na(pre_median_CRE_90days), esrd_kt == 0) %>%
  filter(eGFR_CRE_baseline >= 10)
```

## 6.10 Recalculate baseline eGFR (post-exclusion confirmation)

```{r}
#| label: clean-egfr-recheck
final_master$eGFR_CRE_baseline <- mapply(
  gfr_calc,
  final_master$pre_median_CRE_90days,
  final_master$male,
  final_master$Age
)
```

## 6.11 Define categorical and numeric variables for Table 1

```{r}
#| label: clean-var-lists
all_var <- c(
  "Age", "Gender_Legal_Sex", "male", "Race", "Ethnic_Group",
  ## Outcomes
  "ckd_incidence", "ckd_progression", "eskd_composite", "ckd_composite",
  ## CKD stages
  paste0("ckd_stage_", 1:10, "year"),
  ## Composites by year
  paste0("ckd_composite_", 1:5, "year"),
  ## Comorbidities
  "dm", "htn", "cad", "cirrhosis", "esrd_kt",
  ## Medications
  "ace_arb", "diu", "ppi", "steroids", "smoking", "statins",
  ## ICI / chemo
  "ICI_Class", "nephrotoxic_chemo",
  ## Labs
  "pre_CRE_180days", "pre_HGB_180days", "pre_ALB_180days",
  "eGFR_CRE_baseline", "ckd_stage_baseline",
  ## AKI
  "aki"
)

num_var <- c("Age", "pre_CRE_180days", "pre_HGB_180days", "pre_ALB_180days",
             "eGFR_CRE_baseline")
cat_var <- setdiff(all_var, num_var)

# Only convert variables that exist in the data
cat_var_exist <- intersect(cat_var, names(final_master))
final_master <- final_master %>% mutate(across(all_of(cat_var_exist), factor))

final_master$ICI_Class <- relevel(final_master$ICI_Class, ref = "PD1")
```

---

# 7 Kaplan-Meier Landmark Analysis (Example: 1-Year)

```{r}
#| label: km-1year
test_1year <- final_master %>%
  filter(survival_1year == 1) %>%
  mutate(
    time_to_death_1year_mark = last_cre_days - 365 * 1,
    time_to_death_1year_mark = ifelse(time_to_death_1year_mark > 365 * 9,
                                      365 * 9, time_to_death_1year_mark),
    time_to_death_1year_mark_status = case_when(
      is.na(time_to_death_1year_mark) | time_to_death_1year_mark > 365 * 9 ~ 0,
      !is.na(time_to_death_1year_mark) ~ 1
    )
  )

fit <- survfit(
  Surv(time_to_death_1year_mark, time_to_death_1year_mark_status) ~ ckd_composite_1year,
  data = test_1year
)

ggsurvplot(
  fit, data = test_1year,
  pval = TRUE, pval.coord = c(0, 0.1),
  conf.int = TRUE, risk.table = TRUE,
  xlab = "Days from 1-Year Landmark",
  ylab = "Survival Probability",
  legend.title = "CKD Composite",
  xlim = c(0, 365 * 9), break.time.by = 365,
  ggtheme = theme_minimal()
)
```

---

# 8 Outcome Tables (1–10 Years)

```{r}
#| label: outcome-tables
#| results: asis
for (yr in 1:10) {
  cat(paste0("\n\n## ", yr, "-Year Outcomes\n\n"))
  tbl <- create_outcome_table(final_master, paste0(yr, "year"))
  print(tbl)
}
```

---

# 9 Pipeline Summary

| Section | Step | Key output |
|---|---|---|
| 4.1 | Load master Excel (EMPI + Index_date) | `final_master` base |
| 4.2–4.3 | Load RPDR lab, create `master_lab` | RPDR labs merged with patient index |
| 5.1 | Baseline CRE from RPDR (median ≤ 90 days) | `pre_median_CRE_90days` |
| 5.2 | Baseline eGFR + CKD stage | `eGFR_CRE_baseline`, `ckd_stage_baseline` |
| 5.3 | ESKD composite | `eskd_composite`, `eskd_composite_date` |
| 5.4 | Time-to-event, CKD composite | `time_to_ckd_composite`, `ckd_composite` |
| 5.5 | Survival flags (1–10 yr) | `survival_Xyear` |
| 5.6 | Yearly outcome flags (1–10 yr) | `ckd_incidence_Xyear`, etc. |
| 5.7 | Yearly CRE → eGFR → CKD stage (1–10 yr) | `eGFR_CRE_Xyear`, `ckd_stage_Xyear` |
| 5.8 | Exclusion filters (missing CRE, ESRD, eGFR < 10) | Final cohort |
| 5.10 | Table 1 variable setup | `all_var`, `cat_var`, `num_var` |
| 6 | KM landmark (1-year example) | Survival plot |
| 7 | Outcome tables 1–10 years | `gtsummary` tables |

---

# 10 Input/Variable Summary for RPDR

## 9.1 Required Input Columns (from master Excel)

Your future dataset **must** contain these columns:

| Column | Type | Description |
|---|---|---|
| `EMPI` | character | MRN / Enterprise Master Patient Index |
| `Index_date` | Date | ICI treatment start date (the anchor for all time windows) |
| `last_cre_date` | Date | Date of last creatinine lab (follow-up date) |
| `Age` | numeric | Age at Index_date |
| `Gender_Legal_Sex` | character | |
| `male` | numeric (0/1) | |
| `Race` | character | |
| `Ethnic_Group` | character | |
| `dm` | numeric (0/1) | Diabetes mellitus |
| `htn` | numeric (0/1) | Hypertension |
| `cad` | numeric (0/1) | Coronary artery disease |
| `esrd_kt` | numeric (0/1) | Pre-ICI ESRD / kidney transplant |
| `ace_arb` | numeric (0/1) | ACE inhibitor / ARB use |
| `diu` | numeric (0/1) | Diuretic use |
| `ppi` | numeric (0/1) | PPI use |
| `steroids` | numeric (0/1) | Steroid use |
| `smoking` | numeric (0/1) | Smoking status |
| `Bevacizumab` | numeric (0/1) | Drug indicator |
| `Cisplatin` | numeric (0/1) | Drug indicator |
| `Carboplatin` | numeric (0/1) | Drug indicator |
| `Pemetrexed` | numeric (0/1) | Drug indicator |
| `Gemcitabine` | numeric (0/1) | Drug indicator |
| `Cetuximab` | numeric (0/1) | Drug indicator |
| `Trastuzumab` | numeric (0/1) | Drug indicator |
| `VEGF_TKI` | numeric (0/1) | Drug indicator |
| `pre_CRE_180days` | numeric | Pre-baseline creatinine (180 days, from master) |
| `pre_HGB_180days` | numeric | Pre-baseline hemoglobin (180 days, from master) |
| `pre_ALB_180days` | numeric | Pre-baseline albumin (180 days, from master) |
| `eGFR_cre` | numeric | eGFR from master (pre-computed) |
| `ckd_stage` | numeric | CKD stage from master (pre-computed) |
| `ckd_incidence` | numeric (0/1) | CKD incidence outcome |
| `ckd_incidence_date` | Date | Date of CKD incidence |
| `ckd_progression` | numeric (0/1) | CKD progression outcome |
| `ckd_progression_date` | Date | Date of CKD progression |
| `esrd_kt_after_ici` | numeric (0/1) | ESRD/KT after ICI |
| `esrd_kt_after_ici_date` | Date | Date of ESRD/KT after ICI |
| `eskd_defination_1` | numeric (0/1) | ESKD definition 1 |
| `eskd_defination_1_date` | Date | Date of ESKD definition 1 |
| `death_1year` … `death_10year` | numeric (0/1) | Mortality flags by year |

## 9.2 Required RPDR File

| File | Columns Used | Purpose |
|---|---|---|
| Lab (pipe-delimited `.txt`) | `EMPI`, `Seq_Date_Time`, `Result`, `Group_Id` | Pre-baseline CRE (90d), yearly post-index CRE (1–10 yr) |

The pipeline uses `Group_Id = "CRE"` from the RPDR Lab file.

## 9.3 All Derived Variables (created by this pipeline)

| Variable | Source | Section |
|---|---|---|
| `pre_median_CRE_90days` | RPDR Lab (median CRE ≤ 90d pre-index) | 5.1 |
| `eGFR_CRE_baseline` | CKD-EPI from `pre_median_CRE_90days` | 5.2 |
| `ckd_stage_baseline` | From `eGFR_CRE_baseline` (1–5) | 5.2 |
| `ckd_stage_baseline_3` | 3-level staging | 5.2 |
| `eskd_composite` | esrd_kt_after_ici OR eskd_defination_1 | 5.3 |
| `eskd_composite_date` | pmin of the two ESKD dates | 5.3 |
| `time_to_eskd_composite` | Days from Index_date | 5.4 |
| `time_to_ckd_incidence` | Days from Index_date | 5.4 |
| `time_to_ckd_progression` | Days from Index_date | 5.4 |
| `ckd_composite` | Any of the three outcomes | 5.4 |
| `time_to_ckd_composite` | pmin of the three time-to-event | 5.4 |
| `last_cre_days` | Days from Index_date to last_cre_date | 5.4 |
| `survival_1year` … `survival_10year` | 1 if last_cre_days ≥ 365 × X | 5.5 |
| `ckd_incidence_1year` … `_10year` | 1 if time < 365 × X | 5.6 |
| `ckd_progression_1year` … `_10year` | 1 if time < 365 × X | 5.6 |
| `eskd_composite_1year` … `_10year` | 1 if time < 365 × X | 5.6 |
| `pre_median_CRE_1year` … `_10year` | RPDR Lab (yearly median CRE) | 5.7 |
| `eGFR_CRE_1year` … `_10year` | CKD-EPI from yearly median CRE | 5.7 |
| `ckd_stage_1year` … `_10year` | CKD stage from yearly eGFR | 5.7 |
| `lab_num_1year` … `_10year` | Count of CRE labs in yearly window | 5.7 |
