---
title: "Control Group – Descriptive Analysis"
author: "Tianqi Ouyang"
date: "2026-03-26"
execute:
  echo: false
  include: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = FALSE, include = FALSE)
```

```{r}
#| label: libraries
rm(list = ls())
library(readxl)
library(dplyr)
library(tidyverse)
library(writexl)
library(gtsummary)
library(survival)
library(ggplot2)
library(survminer)
library(ggsurvfit)
library(tidycmprsk)
library(table1)
library(readr)
```

---

::: {style="display:none;"}

# 2 Functions

Reused from `analysis_ici_ckd.qmd`.

## 2.1 pre_lab_median

```{r}
#| label: fn-pre-lab-median
pre_lab_median <- function(data, month, time, index_date, Group_Id) {
  index_date <- enquo(index_date)
  if (month) {
    time_value <- time * 30
  } else {
    time_value <- time
  }

  median_name <- paste0("pre_median_", Group_Id, "_", time_value, "days")

  result <- data %>%
    filter(Group_Id == !!Group_Id, grepl("^\\d+\\.\\d+$", Result)) %>%
    filter(!is.na(Result)) %>%
    mutate(Result = as.numeric(Result)) %>%
    filter(Seq_Date_Time >= (!!index_date) - time_value,
           Seq_Date_Time <= !!index_date) %>%
    group_by(EMPI) %>%
    summarise(
      !!median_name := median(Result, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    select(EMPI, !!median_name)

  return(result)
}
```

## 2.2 pre_lab_median_year

```{r}
#| label: fn-pre-lab-median-year
pre_lab_median_year <- function(data, year, days, index_date, Group_Id) {
  index_date <- enquo(index_date)

  last_offset  <- year * 365
  first_offset <- last_offset - days

  median_name      <- paste0("pre_median_", Group_Id, "_", year, "year")
  lab_num_name     <- paste0("lab_num_", year, "year")
  closest_name     <- paste0("closest_", Group_Id, "_", year, "year")
  closest_date_name <- paste0("closest_", Group_Id, "_", year, "year_date")

  median <- data %>%
    filter(Group_Id == !!Group_Id, grepl("^\\d+\\.\\d+$", Result)) %>%
    filter(!is.na(Result)) %>%
    mutate(Result = as.numeric(Result)) %>%
    filter(Seq_Date_Time >= (!!index_date + first_offset),
           Seq_Date_Time <= !!index_date + last_offset) %>%
    group_by(EMPI) %>%
    summarise(
      !!median_name := median(Result, na.rm = TRUE),
      !!lab_num_name := n(),
      .groups = "drop"
    )

  closest <- data %>%
    filter(Group_Id == !!Group_Id, grepl("^\\d+\\.\\d+$", Result)) %>%
    filter(!is.na(Result)) %>%
    mutate(Result = as.numeric(Result)) %>%
    filter(Seq_Date_Time <= !!index_date + last_offset) %>%
    arrange(EMPI, desc(Seq_Date_Time)) %>%
    distinct(EMPI, .keep_all = TRUE) %>%
    rename(!!closest_name := Result,
           !!closest_date_name := Seq_Date_Time) %>%
    select(EMPI, !!index_date, !!closest_name, !!closest_date_name)

  result <- median %>%
    merge(closest, by = "EMPI", all = TRUE) %>%
    mutate(
      !!median_name := ifelse(is.na(!!sym(median_name)),
                              !!sym(closest_name),
                              !!sym(median_name))
    ) %>%
    select(EMPI, !!median_name, !!lab_num_name)

  return(result)
}
```

## 2.3 left_join_all

```{r}
#| label: fn-left-join-all
left_join_all <- function(master, df_list, by) {
  for (df in df_list) {
    master <- merge(master, df, by = by, all.x = TRUE)
  }
  return(master)
}
```

## 2.4 get_ckd_stage

```{r}
#| label: fn-get-ckd-stage
get_ckd_stage <- function(egfr) {
  if (is.na(egfr)) {
    return(NA)
  } else if (egfr >= 90) {
    return(1)
  } else if (egfr >= 60) {
    return(2)
  } else if (egfr >= 30) {
    return(3)
  } else if (egfr >= 15) {
    return(4)
  } else {
    return(5)
  }
}
```

## 2.5 ckd_epi_gfr (CKD-EPI 2021 creatinine-only)

```{r}
#| label: fn-ckd-epi-gfr
ckd_epi_gfr <- function(creat, male, age) {
  k <- ifelse(male == 1, 0.9, 0.7)
  a <- ifelse(male == 1, -0.302, -0.241)
  m <- ifelse(male == 1, 1, 1.012)
  gfr <- 142 * min(c(creat / k, 1))^a * max(c(creat / k, 1))^-1.2 * 0.9938^age * m
  return(gfr)
}

gfr_calc <- Vectorize(ckd_epi_gfr)
```

---

# 3 Variable Dictionary

| Source (RPDR Dem.txt) | In Code | Derivation |
|---|---|---|
| `EMPI` | `EMPI` | Unique patients from Dem.txt |
| First neoplasm Dx date (Dia.txt) | `Index_date` | First ICD10 C00–D49 diagnosis date |
| `Date_of_Birth` | `Date_of_Birth` | Used to derive `Age` |
| `Gender_Legal_Sex` | `Gender_Legal_Sex` | Used to derive `male` |
| `Race1` | `Race1` | As-is |
| `Ethnic_Group` | `Ethnic_Group` | As-is |
| `Date_Of_Death` | `Date_Of_Death` | Used to derive `death`, `time_to_death` |
| Last CRE date (Lab.txt) | `last_cre_date` | Derived: last CRE lab date per patient |

### Derived variables

| Variable | Description |
|---|---|
| `Age` | `floor((Index_date − Date_of_Birth) / 365.25)` |
| `male` | 1 if Male, else 0 |
| `last_cre_days` | Days from `Index_date` to `last_cre_date` |
| `death` | 1 if `Date_Of_Death` is not NA |
| `time_to_death` | `Date_Of_Death − Index_date` (days), or `last_cre_days` if alive |
| `death_Xyear` (1–10) | 1 if died within X×365 days |
| `survival_Xyear` (1–10) | 1 if `last_cre_days ≥ 365×X` |
| `pre_median_CRE_90days` | Median CRE ≤90 days pre-index (RPDR) |
| `eGFR_CRE_baseline` | CKD-EPI 2021 from baseline CRE |
| `ckd_stage_baseline` | CKD stages 1–5 |
| `ckd_stage_baseline_3` | 3-level: stage 1 / stage 2 / stage 3 or more |
| `pre_median_CRE_Xyear` (1–10) | Yearly median CRE (90-day window) |
| `eGFR_CRE_Xyear` (1–10) | Yearly eGFR |
| `ckd_stage_Xyear` (1–10) | Yearly CKD stage |
| `lab_num_Xyear` (1–10) | Number of CRE labs in yearly window |

#### Matching variables (CRE measurement counts)

| Variable | Description |
|---|---|
| `n_cre_pre_1yr` | Count of CRE labs in `[Index_date − 365, Index_date)` — primary matching variable |
| `n_cre_pre_all` | Count of all CRE labs before `Index_date` |
| `pre_history_days` | Days from earliest pre-index CRE lab to `Index_date` (0 if none) |
| `cre_rate_pre` | `n_cre_pre_all / (pre_history_days / 365)` — NA if no pre-index history |
| `has_pre_cre` | 1 if `n_cre_pre_1yr > 0`, else 0 |
| `n_cre_post` | Count of CRE labs in `(Index_date, last_cre_date]` (covariate, not for matching) |
| `cre_rate_post` | `n_cre_post / (last_cre_days / 365)` — NA if `last_cre_days ≤ 0` |

### Inclusion criteria

| Criterion | Description |
|---|---|
| Baseline CRE | ≥1 creatinine within 12 months prior to Index_date |
| Follow-up | ≥18 months (548 days) of follow-up from Index_date |
| Post-index CRE | ≥2 creatinine results after Index_date |

> Follow-up uses `last_cre_date` → `last_cre_days` for censoring.

---

# 4 Data Loading

## 4.1 Load RPDR files (Dem, Lab from 04092026 pull; Dia, Med from original pull)

```{r}
#| label: load-rpdr
rpdr_path_new <- "/Users/to909/Partners HealthCare Dropbox/Tianqi Ouyang/CKD after ICI/RPDR/RPDR Joyita/04092026/"
rpdr_path_old <- "/Users/to909/Partners HealthCare Dropbox/Tianqi Ouyang/CKD after ICI/RPDR/RPDR Joyita/"

# Demographics — cohort source (new pull)
dem <- read.table(
  paste0(rpdr_path_new, "mes98_040726114217167239_Dem.txt"),
  sep = "|", fill = TRUE, quote = "", header = TRUE
)
dem$Date_of_Birth <- as.Date(dem$Date_of_Birth, "%m/%d/%Y")
dem$Date_Of_Death <- as.Date(dem$Date_Of_Death, "%m/%d/%Y")

# Lab (new pull)
lab <- read.table(
  paste0(rpdr_path_new, "mes98_040726114217167239_Lab.txt"),
  sep = "|", fill = TRUE, quote = "", header = TRUE
) %>%
  select(EMPI, EPIC_PMRN, Seq_Date_Time, Result, Group_Id, Test_Description)
lab$Seq_Date_Time <- as.Date(lab$Seq_Date_Time, "%m/%d/%Y")

# Diagnoses (old pull — Dia.txt not included in 04092026)
dia <- read.table(
  paste0(rpdr_path_old, "MES98_20251202_101837_Dia.txt"),
  sep = "|", fill = TRUE, quote = "", header = TRUE
) %>%
  select(EMPI, EPIC_PMRN, Code, Date, Diagnosis_Name)
dia$Date <- as.Date(dia$Date, "%m/%d/%Y")

# Medications (old pull — Med.txt not included in 04092026)
med <- read.table(
  paste0(rpdr_path_old, "MES98_20251202_101837_Med.txt"),
  sep = "|", fill = TRUE, quote = "", header = TRUE
) %>%
  select(EMPI, Medication_Date, Medication)
med$Medication_Date <- as.Date(med$Medication_Date, "%m/%d/%Y")
```

## 4.2 Define cohort from Dem.txt

```{r}
#| label: define-cohort
control <- dem %>%
  distinct(EMPI, .keep_all = TRUE) %>%
  select(EMPI, Date_of_Birth, Date_Of_Death, Gender_Legal_Sex, Race1, Ethnic_Group)
cat("Unique patients from Dem.txt:", nrow(control), "\n")
```

## 4.3 Derive Index_date from first neoplasm diagnosis (ICD10 C00–D49)

```{r}
#| label: derive-index-date
# First diagnosis with ICD10 code C00–D49 (all neoplasms)
neoplasm_index <- dia %>%
  filter(grepl("^C|^D[0-4]", Code, ignore.case = TRUE)) %>%
  arrange(EMPI, Date) %>%
  distinct(EMPI, .keep_all = TRUE) %>%
  select(EMPI, Index_date = Date)

control <- control %>%
  merge(neoplasm_index, by = "EMPI") %>%
  filter(!is.na(Index_date))

cat("Patients with neoplasm diagnosis (Index_date):", nrow(control), "\n")
```

## 4.4 Derive last_cre_date from RPDR Lab

```{r}
#| label: derive-last-cre
last_cre <- lab %>%
  filter(Group_Id == "CRE", grepl("^\\d+\\.\\d+$", Result)) %>%
  filter(!is.na(Result)) %>%
  arrange(EMPI, desc(Seq_Date_Time)) %>%
  distinct(EMPI, .keep_all = TRUE) %>%
  select(EMPI, last_cre_date = Seq_Date_Time)

control <- control %>%
  merge(last_cre, by = "EMPI", all.x = TRUE)
```

## 4.5 Derive variables

```{r}
#| label: derive-vars
control <- control %>%
  mutate(
    group         = "control",
    Index_date    = as.Date(Index_date),
    Date_of_Birth = as.Date(Date_of_Birth),
    Date_Of_Death = as.Date(Date_Of_Death),
    last_cre_date = as.Date(last_cre_date),
    Age           = as.numeric(floor(difftime(Index_date, Date_of_Birth, units = "days") / 365.25)),
    male          = ifelse(Gender_Legal_Sex == "Male", 1, 0),
    last_cre_days = as.numeric(difftime(last_cre_date, Index_date, units = "days")),
    death         = ifelse(!is.na(Date_Of_Death), 1, 0),
    death_days    = as.numeric(difftime(Date_Of_Death, Index_date, units = "days")),
    time_to_death = death_days,
    time_to_death_status = ifelse(!is.na(death_days), 2, 0),
    Race1 = ifelse(Race1 == "White", "White",
            ifelse(Race1 == "Black", "Black",
            ifelse(Race1 == "Asian", "Asian", "Other/Unknown"))),
    Ethnic_Group = case_when(
      Ethnic_Group == "HISPANIC" ~ "Hispanic",
      Ethnic_Group == "Non Hispanic" ~ "Non_hispanic",
      TRUE ~ "Other"
    )
  ) %>%
  filter(last_cre_days > 0)
```

## 4.6 Inclusion criteria

```{r}
#| label: inclusion-criteria
n_before <- nrow(control)

# 1. At least 1 creatinine within 12 months (365 days) prior to Index_date
cre_within_12mo <- lab %>%
  filter(Group_Id == "CRE", grepl("^\\d+\\.\\d+$", Result), !is.na(Result)) %>%
  merge(control %>% select(EMPI, Index_date), by = "EMPI") %>%
  filter(Seq_Date_Time >= Index_date - 365, Seq_Date_Time <= Index_date) %>%
  distinct(EMPI)

control <- control %>% filter(EMPI %in% cre_within_12mo$EMPI)
cat("After baseline CRE within 12mo:", nrow(control), "of", n_before, "\n")

# 2. At least 18 months (548 days) follow-up
n_before_fu <- nrow(control)
control <- control %>% filter(last_cre_days >= 548)
cat("After 18mo follow-up:", nrow(control), "of", n_before_fu, "\n")

# 3. At least 2 creatinine results after Index_date
n_before_cre2 <- nrow(control)
post_cre_count <- lab %>%
  filter(Group_Id == "CRE", grepl("^\\d+\\.\\d+$", Result), !is.na(Result)) %>%
  merge(control %>% select(EMPI, Index_date), by = "EMPI") %>%
  filter(Seq_Date_Time > Index_date) %>%
  group_by(EMPI) %>%
  summarise(n_post_cre = n(), .groups = "drop") %>%
  filter(n_post_cre >= 2)

control <- control %>% filter(EMPI %in% post_cre_count$EMPI)
cat("After >=2 post-index CRE:", nrow(control), "of", n_before_cre2, "\n")
cat("Final control cohort:", nrow(control), "\n")
```

## 4.7 Death flags (1–10 years)

Uses `last_date` logic from Kavita master builder: `last_date = if_else(is.na(Date_Of_Death), last_cre_date, Date_Of_Death)`.

```{r}
#| label: death-flags
control <- control %>%
  mutate(
    last_date = if_else(is.na(Date_Of_Death), last_cre_date, Date_Of_Death),
    last_days = as.numeric(last_date - Index_date)
  )

for (yr in 1:10) {
  control[[paste0("death_", yr, "year")]] <-
    ifelse(!is.na(control$death_days) & control$death_days < 365 * yr, 1, 0)
}
```

## 4.8 Create master_lab / master_dia / master_med

```{r}
#| label: create-master
master_list <- control %>% select(EMPI, Index_date)
master_lab  <- lab %>% merge(master_list, by = "EMPI")
master_dia  <- dia %>% merge(master_list, by = "EMPI")
master_med  <- med %>% merge(master_list, by = "EMPI")
```

## 4.9 RPDR-derived functions

```{r}
#| label: fn-rpdr-helpers
med_list <- function(names, index_date, excluded_names = NULL, data) {
  if (!is.null(excluded_names)) {
    result <- data %>%
      filter(Medication_Date <= {{index_date}}) %>%
      filter(
        grepl(paste(names, collapse = "|"), Medication, ignore.case = TRUE),
        !grepl(paste(excluded_names, collapse = "|"), Medication, ignore.case = TRUE)
      ) %>%
      distinct(EMPI)
  } else {
    result <- data %>%
      filter(Medication_Date <= {{index_date}}) %>%
      filter(grepl(paste(names, collapse = "|"), Medication, ignore.case = TRUE)) %>%
      distinct(EMPI)
  }
  return(result)
}

dia_list_icd <- function(data, index_date, icd_code, excluded_names = NULL,
                         med_data = NULL, med_names = NULL) {
  if (is.null(excluded_names)) {
    dia_result <- data %>%
      filter(Date <= {{index_date}}) %>%
      filter(grepl(paste(icd_code, collapse = "|"), Code, ignore.case = TRUE)) %>%
      distinct(EMPI)
  } else {
    dia_result <- data %>%
      filter(Date <= {{index_date}}) %>%
      filter(
        grepl(paste(icd_code, collapse = "|"), Code, ignore.case = TRUE),
        !grepl(paste(excluded_names, collapse = "|"), Diagnosis_Name, ignore.case = TRUE)
      ) %>%
      distinct(EMPI)
  }
  if (!is.null(med_data)) {
    med_result <- med_data %>%
      filter(Medication_Date <= {{index_date}}) %>%
      filter(grepl(paste(med_names, collapse = "|"), Medication, ignore.case = TRUE)) %>%
      distinct(EMPI)
  } else {
    med_result <- NULL
  }
  rbind(dia_result, med_result) %>% distinct(EMPI)
}

dia_list_name <- function(data, listed_names, index_date, excluded_names = NULL,
                          med_data = NULL, med_names = NULL) {
  if (is.null(excluded_names)) {
    dia_result <- data %>%
      filter(Date <= {{index_date}}) %>%
      filter(grepl(paste(listed_names, collapse = "|"), Diagnosis_Name, ignore.case = TRUE)) %>%
      distinct(EMPI)
  } else {
    excluded_pattern <- paste(gsub("([][()])", "\\\\\\1", excluded_names), collapse = "|")
    dia_result <- data %>%
      filter(Date <= {{index_date}}) %>%
      filter(
        grepl(paste(listed_names, collapse = "|"), Diagnosis_Name, ignore.case = TRUE),
        !grepl(excluded_pattern, Diagnosis_Name, ignore.case = TRUE)
      ) %>%
      distinct(EMPI)
  }
  if (!is.null(med_data)) {
    med_result <- med_data %>%
      filter(Medication_Date <= {{index_date}}) %>%
      filter(grepl(paste(med_names, collapse = "|"), Medication, ignore.case = TRUE)) %>%
      distinct(EMPI)
  } else {
    med_result <- NULL
  }
  rbind(dia_result, med_result) %>% distinct(EMPI)
}

dia_list_icd_after_index <- function(data, index_date, icd_code, excluded_names = NULL) {
  if (is.null(excluded_names)) {
    result <- data %>%
      filter(Date > {{index_date}}) %>%
      filter(grepl(paste(icd_code, collapse = "|"), Code, ignore.case = TRUE))
  } else {
    result <- data %>%
      filter(Date > {{index_date}}) %>%
      filter(
        grepl(paste(icd_code, collapse = "|"), Code, ignore.case = TRUE),
        !grepl(paste(excluded_names, collapse = "|"), Diagnosis_Name, ignore.case = TRUE)
      )
  }
  result %>%
    arrange(EMPI, Date) %>%
    distinct(EMPI, .keep_all = TRUE) %>%
    select(EMPI, first_event_date = Date)
}

make_ckd_outcome_date <- function(data, id_var, date_var, condition_expr,
                                   days_threshold = 90, outcome_name = "default") {
  id_var      <- enquo(id_var)
  date_var    <- enquo(date_var)
  cond_q      <- enquo(condition_expr)
  outcome_sym <- ensym(outcome_name)
  outcome_date_name <- paste0(outcome_name, "_date")

  working <- data %>%
    arrange(!!id_var, !!date_var) %>%
    group_by(!!id_var) %>%
    mutate(
      .cond   = if_else(!!cond_q, 1L, 0L),
      .change = .cond != lag(.cond, default = first(.cond)),
      .run    = cumsum(.change)
    )

  episodes <- working %>%
    group_by(!!id_var, .run) %>%
    summarise(
      .code = first(.cond), .start = first(!!date_var),
      .last = last(!!date_var), .groups = "drop"
    ) %>%
    group_by(!!id_var) %>%
    mutate(.end = lead(.start),
           .end = if_else(is.na(.end), .last, .end),
           .days = as.integer(.end - .start)) %>%
    ungroup()

  episodes %>%
    filter(!is.na(.days)) %>%
    group_by(!!id_var) %>%
    summarise(
      has_event  = any(.code == 1L & .days >= days_threshold, na.rm = TRUE),
      event_date = if (any(.code == 1L & .days >= days_threshold, na.rm = TRUE)) {
        min(.start[.code == 1L & .days >= days_threshold], na.rm = TRUE)
      } else { as.Date(NA) },
      .groups = "drop"
    ) %>%
    mutate(!!outcome_sym := as.integer(has_event)) %>%
    rename(!!outcome_date_name := event_date) %>%
    select(-has_event)
}
```

## 4.10 Derive comorbidities & medications from RPDR

```{r}
#| label: derive-comorbidities
# Medication name lists
ppi_names <- c("omeprazole", "Prilosec", "lansoprazole", "Prevacid",
               "dexlansoprazole", "rabeprazole", "Aciphex",
               "pantoprazole", "Protonix", "esomeprazole", "Nexium")
ace_arb_names <- c("benazepril", "captopril", "enalapril", "fosinopril",
                   "lisinopril", "ramipril", "losartan", "valsartan",
                   "irbesartan", "olmesartan", "candesartan", "telmisartan",
                   "azilsartan", "eprosartan")
diuretics_names <- c("bumetanide", "furosemide", "torsemide",
                     "hydrochlorothiazide", "chlorthalidone", "metolazone",
                     "spironolactone", "triamterene", "amiloride", "eplerenone")
steroids_names <- c("prednisone", "methylprednisolone", "prednisolone",
                    "dexamethasone", "budesonide")
no_steroid_names <- c("Inhaler", "Nebulization", "Ointment", "Ophthalmic")
statins_names <- c("Atorvastatin", "Lipitor", "Fluvastatin", "Lovastatin",
                   "Pitavastatin", "Pravastatin", "Rosuvastatin", "Simvastatin")
dpp4i_names <- c("Alogliptin", "Vipidia", "Nesina",
                 "Linagliptin", "Tradjenta",
                 "Saxagliptin", "Onglyza",
                 "Sitagliptin", "Januvia",
                 "Vildagliptin", "Galvus")
sglt2i_names <- c("Canagliflozin", "Dapagliflozin", "Empagliflozin", "Ertugliflozin",
                  "Invokana", "Farxiga", "Forxiga", "Jardiance", "Steglatro")

# Diagnosis codes
diabetes_codes <- c("250.00", "250.01", "250.02", "250.03", "E10.9", "E11.9",
                    "E10.65", "E11.65", "E10.21", "E11.21", "E10.22", "E11.22",
                    "E13.9")
htn_codes <- c("401.1", "401.0", "401.9", "I10", "I15.1", "I15.9")
no_htn_list <- c("Pulmonary Embolism", "Borderline hypertension",
                 "Chronic venous hypertension")
htn_med_list <- c("hydrochlorothiazide", "amlodipine", "lisinopril",
                  "losartan", "valsartan", "diltiazem", "clonidine")
cad_list <- c("Coronary artery disease", "Coronary atherosclerosis",
              "Atherosclerotic heart disease", "Angina pectoris",
              "Unstable angina")
smoking_list <- c("smoking", "tobacco", "cigarette", "nicotine", "smoke")
esrd_kt_codes <- c("585.6", "N18.6", "V45.1", "Z94.0", "Z99.2",
                   "996.81", "T86.1", "T86.10", "T86.11", "T86.12")

# Derive
ppi      <- med_list(data = master_med, names = ppi_names, index_date = Index_date)
ace_arb  <- med_list(data = master_med, names = ace_arb_names, index_date = Index_date)
diu      <- med_list(data = master_med, names = diuretics_names, index_date = Index_date)
steroids <- med_list(data = master_med, names = steroids_names,
                     index_date = Index_date, excluded_names = no_steroid_names)
statins  <- med_list(data = master_med, names = statins_names, index_date = Index_date)
dpp4i    <- med_list(data = master_med, names = dpp4i_names, index_date = Index_date)
sglt2i   <- med_list(data = master_med, names = sglt2i_names, index_date = Index_date)
esrd_kt  <- dia_list_icd(master_dia, icd_code = esrd_kt_codes, index_date = Index_date)
dm  <- dia_list_icd(master_dia, icd_code = diabetes_codes, index_date = Index_date)
htn <- dia_list_icd(master_dia, icd_code = htn_codes, excluded_names = no_htn_list,
                    med_data = master_med, med_names = htn_med_list,
                    index_date = Index_date)
cad <- dia_list_name(master_dia, listed_names = cad_list, index_date = Index_date)
cirrhosis <- master_dia %>%
  filter(Date <= Index_date) %>%
  filter(grepl("cirrhosis", Diagnosis_Name, ignore.case = TRUE)) %>%
  distinct(EMPI)
smoking_dx <- master_dia %>%
  filter(grepl(paste(smoking_list, collapse = "|"), Diagnosis_Name, ignore.case = TRUE)) %>%
  distinct(EMPI)

control <- control %>%
  mutate(
    diu       = ifelse(EMPI %in% diu$EMPI, 1, 0),
    ace_arb   = ifelse(EMPI %in% ace_arb$EMPI, 1, 0),
    esrd_kt   = ifelse(EMPI %in% esrd_kt$EMPI, 1, 0),
    dm        = ifelse(EMPI %in% dm$EMPI, 1, 0),
    htn       = ifelse(EMPI %in% htn$EMPI, 1, 0),
    ppi       = ifelse(EMPI %in% ppi$EMPI, 1, 0),
    steroids  = ifelse(EMPI %in% steroids$EMPI, 1, 0),
    smoking   = ifelse(EMPI %in% smoking_dx$EMPI, 1, 0),
    cad       = ifelse(EMPI %in% cad$EMPI, 1, 0),
    cirrhosis = ifelse(EMPI %in% cirrhosis$EMPI, 1, 0),
    statins   = ifelse(EMPI %in% statins$EMPI, 1, 0),
    dpp4i     = ifelse(EMPI %in% dpp4i$EMPI, 1, 0),
    sglt2i    = ifelse(EMPI %in% sglt2i$EMPI, 1, 0)
  )
```

---

# 5 Baseline Lab Values

## 5.1 Baseline creatinine (median within 90 days pre-index)

```{r}
#| label: baseline-cre
pre_lab_baseline <- pre_lab_median(
  master_lab, month = FALSE, time = 90,
  index_date = Index_date, Group_Id = "CRE"
)
pre_lab_hgb <- pre_lab_median(
  master_lab, month = FALSE, time = 90,
  index_date = Index_date, Group_Id = "HGB"
)
pre_lab_alb <- pre_lab_median(
  master_lab, month = FALSE, time = 90,
  index_date = Index_date, Group_Id = "ALB"
)
control <- control %>%
  merge(pre_lab_baseline, by = "EMPI", all.x = TRUE) %>%
  merge(pre_lab_hgb, by = "EMPI", all.x = TRUE) %>%
  merge(pre_lab_alb, by = "EMPI", all.x = TRUE)
```

## 5.1b CRE measurement counts (for matching)

```{r}
#| label: cre-measurement-counts
# CRE labs restricted to valid numeric results, tagged with Index_date
cre_labs <- master_lab %>%
  filter(Group_Id == "CRE", grepl("^\\d+\\.\\d+$", Result)) %>%
  select(EMPI, Seq_Date_Time, Index_date)

# Pre-index counts and rate
pre_cre_counts <- cre_labs %>%
  filter(Seq_Date_Time < Index_date) %>%
  group_by(EMPI) %>%
  summarise(
    n_cre_pre_all    = n(),
    n_cre_pre_1yr    = sum(Seq_Date_Time >= Index_date - 365),
    pre_history_days = as.numeric(Index_date[1] - min(Seq_Date_Time)),
    .groups = "drop"
  ) %>%
  mutate(
    cre_rate_pre = ifelse(pre_history_days > 0,
                          n_cre_pre_all / (pre_history_days / 365),
                          NA_real_),
    has_pre_cre  = as.integer(n_cre_pre_1yr > 0)
  )

# Post-index count, bounded by last_cre_date
post_cre_counts <- cre_labs %>%
  inner_join(control %>% select(EMPI, last_cre_date), by = "EMPI") %>%
  filter(Seq_Date_Time > Index_date, Seq_Date_Time <= last_cre_date) %>%
  group_by(EMPI) %>%
  summarise(n_cre_post = n(), .groups = "drop")

control <- control %>%
  merge(pre_cre_counts,  by = "EMPI", all.x = TRUE) %>%
  merge(post_cre_counts, by = "EMPI", all.x = TRUE) %>%
  mutate(
    n_cre_pre_all    = replace_na(n_cre_pre_all, 0),
    n_cre_pre_1yr    = replace_na(n_cre_pre_1yr, 0),
    pre_history_days = replace_na(pre_history_days, 0),
    has_pre_cre      = replace_na(has_pre_cre, 0),
    n_cre_post       = replace_na(n_cre_post, 0),
    cre_rate_post    = ifelse(last_cre_days > 0,
                              n_cre_post / (last_cre_days / 365),
                              NA_real_)
  )
```

## 5.2 Baseline eGFR & CKD stage

```{r}
#| label: baseline-egfr
control$eGFR_CRE_baseline <- mapply(
  gfr_calc,
  control$pre_median_CRE_90days,
  control$male,
  control$Age
)

control <- control %>%
  mutate(
    ckd_stage_baseline = case_when(
      eGFR_CRE_baseline >= 90                           ~ 1,
      eGFR_CRE_baseline >= 60 & eGFR_CRE_baseline < 90 ~ 2,
      eGFR_CRE_baseline >= 30 & eGFR_CRE_baseline < 60 ~ 3,
      eGFR_CRE_baseline >= 15 & eGFR_CRE_baseline < 30 ~ 4,
      eGFR_CRE_baseline < 15                            ~ 5
    ),
    ckd_stage_baseline_3 = case_when(
      ckd_stage_baseline == 1  ~ "stage 1",
      ckd_stage_baseline == 2  ~ "stage 2",
      ckd_stage_baseline >= 3  ~ "stage 3 or more"
    )
  )
```

## 5.3 CKD outcomes from RPDR Lab (sustained eGFR-based)

```{r}
#| label: clean-ckd-outcomes
outcome_labs <- master_lab %>%
  select(EMPI, Seq_Date_Time, Result, Group_Id, Index_date) %>%
  merge(control %>% select(EMPI, pre_median_CRE_90days, eGFR_CRE_baseline,
                            male, Age),
        by = "EMPI") %>%
  filter(Group_Id == "CRE", grepl("^\\d+\\.\\d+$", Result)) %>%
  filter(!is.na(Result)) %>%
  mutate(Result = as.numeric(Result)) %>%
  filter(Seq_Date_Time > Index_date) %>%
  mutate(eGFR_CRE = mapply(gfr_calc, Result, male, Age)) %>%
  select(EMPI, eGFR_CRE_baseline, eGFR_CRE, eGFR_CRE_date = Seq_Date_Time) %>%
  mutate(pct_drop = round((eGFR_CRE_baseline - eGFR_CRE) / eGFR_CRE_baseline, 3)) %>%
  arrange(EMPI, eGFR_CRE_date)

ckd_incidence <- make_ckd_outcome_date(
  outcome_labs, EMPI, eGFR_CRE_date,
  pct_drop > 0.25 & eGFR_CRE < 60 & eGFR_CRE_baseline >= 60,
  90, "ckd_incidence")

eskd_defination_1 <- make_ckd_outcome_date(
  outcome_labs, EMPI, eGFR_CRE_date,
  eGFR_CRE < 10, 90, "eskd_defination_1")

ckd_progression <- make_ckd_outcome_date(
  outcome_labs, EMPI, eGFR_CRE_date,
  pct_drop > 0.30 & eGFR_CRE_baseline < 60, 90, "ckd_progression")

control <- control %>%
  merge(ckd_incidence, by = "EMPI", all.x = TRUE) %>%
  merge(eskd_defination_1, by = "EMPI", all.x = TRUE) %>%
  merge(ckd_progression, by = "EMPI", all.x = TRUE) %>%
  mutate(
    ckd_incidence     = replace_na(ckd_incidence, 0),
    eskd_defination_1 = replace_na(eskd_defination_1, 0),
    ckd_progression   = replace_na(ckd_progression, 0),
    esrd_kt_after_ici = ifelse(EMPI %in% (dia_list_icd_after_index(
      master_dia, icd_code = esrd_kt_codes, index_date = Index_date)$EMPI), 1, 0),
    eskd_composite    = ifelse(esrd_kt_after_ici == 1 | eskd_defination_1 == 1, 1, 0),
    ckd_composite     = ifelse(eskd_composite == 1 | ckd_incidence == 1 | ckd_progression == 1, 1, 0),
    time_to_ckd_incidence   = as.numeric(ckd_incidence_date - Index_date),
    time_to_ckd_progression = as.numeric(ckd_progression_date - Index_date),
    time_to_eskd_composite  = as.numeric(pmin(eskd_defination_1_date, na.rm = TRUE) - Index_date),
    time_to_ckd_composite   = pmin(time_to_ckd_incidence, time_to_ckd_progression,
                                    time_to_eskd_composite, na.rm = TRUE)
  )
```

---

# 6 Yearly Lab Values & Survival Flags

## 6.1 Survival flags (1–10 years)

```{r}
#| label: survival-flags
for (yr in 1:10) {
  control[[paste0("survival_", yr, "year")]] <-
    ifelse(control$last_cre_days >= 365 * yr, 1, 0)
}
```

## 6.2 Yearly CRE medians, eGFR, CKD stage (1–10 years)

```{r}
#| label: yearly-cre
yearly_cre_list <- list()
for (yr in 1:10) {
  yearly_cre_list[[yr]] <- pre_lab_median_year(
    master_lab, year = yr, days = 90,
    index_date = Index_date, Group_Id = "CRE"
  )
}

control <- left_join_all(control, yearly_cre_list, by = "EMPI")

for (yr in 1:10) {
  cre_col   <- paste0("pre_median_CRE_", yr, "year")
  egfr_col  <- paste0("eGFR_CRE_", yr, "year")
  stage_col <- paste0("ckd_stage_", yr, "year")

  control[[egfr_col]]  <- mapply(gfr_calc, control[[cre_col]],
                                  control$male, control$Age)
  control[[stage_col]] <- sapply(control[[egfr_col]], get_ckd_stage)
}
```

---

## 6.3 AKI (from RPDR Lab)

```{r}
#| label: clean-aki
aki_data <- master_lab %>%
  merge(pre_lab_baseline, by = "EMPI", all.x = TRUE) %>%
  filter(Group_Id == "CRE", grepl("^\\d+\\.\\d+$", Result)) %>%
  filter(!is.na(Result), !is.na(pre_median_CRE_90days)) %>%
  mutate(Result = as.numeric(Result)) %>%
  filter(Seq_Date_Time > Index_date,
         Seq_Date_Time <= Index_date + 90) %>%
  filter(Result > 1.5 * pre_median_CRE_90days) %>%
  arrange(EMPI, Seq_Date_Time) %>%
  distinct(EMPI, .keep_all = TRUE) %>%
  select(EMPI, aki_date = Seq_Date_Time)

control <- control %>%
  merge(aki_data, by = "EMPI", all.x = TRUE) %>%
  mutate(aki = ifelse(!is.na(aki_date), 1, 0))
```

## 6.4 Competing risk & analysis variables

```{r}
#| label: competing-risk-vars
control <- control %>%
  mutate(
    year_10 = Age / 10,
    eGFR_CRE_baseline_decrease_10 = (eGFR_CRE_baseline * -1) / 10,
    # Competing risk: pmin of all event times, then priority-based status
    time_to_ckd_cmp = pmin(time_to_ckd_composite, last_cre_days, death_days, na.rm = TRUE),
    status_ckd_cmp = case_when(
      # Priority 1: Death (2)
      !is.na(death_days) & time_to_ckd_cmp == death_days ~ 2,
      # Priority 2: CKD Event (1)
      !is.na(time_to_ckd_composite) & time_to_ckd_cmp == time_to_ckd_composite ~ 1,
      # Priority 3: Censored / Last Creatinine (0)
      !is.na(last_cre_days) & time_to_ckd_cmp == last_cre_days ~ 0,
      TRUE ~ NA_real_
    ),
    status_ckd_cmp = factor(status_ckd_cmp, levels = c(0, 1, 2),
                             labels = c("Censored", "CKD Composite", "Death")),
    aki_before_ckd = ifelse(as.numeric(aki_date - Index_date) <= time_to_ckd_composite, 1, 0)
  )

# Yearly outcome flags
for (yr in 1:10) {
  control[[paste0("ckd_incidence_", yr, "year")]] <-
    ifelse(control$time_to_ckd_incidence < 365 * yr, 1, 0)
  control[[paste0("ckd_progression_", yr, "year")]] <-
    ifelse(control$time_to_ckd_progression < 365 * yr, 1, 0)
  control[[paste0("eskd_composite_", yr, "year")]] <-
    ifelse(control$time_to_eskd_composite < 365 * yr, 1, 0)
  control[[paste0("ckd_composite_", yr, "year")]] <-
    ifelse(control$time_to_ckd_composite < 365 * yr, 1, 0)
}

control <- control %>%
  mutate(across(
    .cols = matches("(ckd_incidence|ckd_progression|eskd_composite|ckd_composite).*year$"),
    .fns  = ~ factor(replace_na(., 0))
  ))
```

---

# 7 Exclusion & Final Dataset

## 7.1 Exclude missing baseline CRE and eGFR < 10

```{r}
#| label: exclusion
n_before <- nrow(control)

control <- control %>%
  filter(!is.na(pre_median_CRE_90days)) %>%
  filter(eGFR_CRE_baseline >= 10)

n_after <- nrow(control)
cat("Excluded", n_before - n_after, "patients.",
    n_after, "patients remain.\n")
```

## 7.2 Save control_group.xlsx

```{r}
#| label: save-xlsx
write_xlsx(control, "control_group.xlsx")
cat("Saved control_group.xlsx with", nrow(control), "rows and",
    ncol(control), "columns.\n")
```

---

:::

# Descriptive Analysis {.unnumbered}

## Index Year Distribution {.unnumbered}

```{r}
#| label: index-year-histogram
#| include: true
#| fig-cap: "Distribution of Index_date (first neoplasm diagnosis) by year"
control %>%
  mutate(index_year = as.numeric(format(Index_date, "%Y"))) %>%
  ggplot(aes(x = index_year)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "white") +
  labs(x = "Year of Cancer Diagnosis", y = "Count",
       title = "Distribution of Index Year – Control Group") +
  theme_minimal()
```

## Table 1 – Demographics {.unnumbered}

```{r}
#| label: table1-demographics
#| include: true
control %>%
  select(
    Age, Gender_Legal_Sex, Race1, Ethnic_Group,
    ## Comorbidities
    dm, htn, cad, cirrhosis,
    ## Medications
    ace_arb, diu, ppi, steroids, smoking, statins, dpp4i, sglt2i,
    ## Baseline Labs
    eGFR_CRE_baseline, ckd_stage_baseline,
    pre_median_HGB_90days, pre_median_ALB_90days,
    ## Matching — CRE surveillance
    n_cre_pre_1yr, n_cre_pre_all, pre_history_days, cre_rate_pre, has_pre_cre,
    n_cre_post, cre_rate_post,
    ## AKI
    aki
  ) %>%
  tbl_summary(
    type = list(Age ~ "continuous", eGFR_CRE_baseline ~ "continuous",
                pre_median_HGB_90days ~ "continuous",
                pre_median_ALB_90days ~ "continuous",
                n_cre_pre_1yr ~ "continuous",
                n_cre_pre_all ~ "continuous",
                pre_history_days ~ "continuous",
                cre_rate_pre ~ "continuous",
                n_cre_post ~ "continuous",
                cre_rate_post ~ "continuous"),
    statistic = list(all_continuous() ~ "{median} ({p25}, {p75})"),
    missing_text = "Missing"
  ) %>%
  bold_labels()
```
