---
title: "Propensity-Score Matching – Treated vs Control"
author: "Tianqi Ouyang"
date: "2026-04-15"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
execute:
  warning: false
  message: false
---

# Summary

## Rationale

Treated (ICI, n=379) and control (non-ICI neoplasm, n=1,307) cohorts differ on demographics, calendar era, comorbidity profile, baseline kidney function, and — critically for outcome interpretation — **baseline creatinine surveillance intensity** (`n_cre_pre_1yr`). A direct comparison would confound the ICI effect with differential measurement, follow-up, and practice-era effects. Propensity-score matching is used to construct a comparable control subset while retaining all 379 treated patients.

## Method

| Parameter | Value |
|---|---|
| Matching algorithm | Nearest-neighbor PS matching (logit PS), greedy, no replacement |
| Ratio | Variable 1:k, `ratio = 2` (target mean), `min.controls = 1`, `max.controls = 3` |
| Caliper | 0.2 × SD(logit PS) |
| Exact-match stratum | `era` (2016–2019 / 2020–2023) |
| Eligibility filter | `has_pre_cre == 1` |
| Seed | `set.seed(42)` |
| Software | R `MatchIt` 4.x |
| Balance target | Standardized mean difference (SMD) < 0.1 for all PS model covariates |

## How the control cohort reaches n = 1,307 (before matching)

The 04092026 RPDR pull is **pre-curated** — the source `Dem.txt` already contains exactly 1,307 patients who satisfy every cohort-build inclusion criterion. The control pipeline filters re-confirm but do not drop additional rows.

| Step (executed in `control_group.qmd`) | n |
|---|---|
| 1. Unique EMPIs in `mes98_040726114217167239_Dem.txt` | 1,307 |
| 2. Has neoplasm Dx (ICD10 C00–D49) → defines `Index_date` | 1,307 |
| 3. Has post-index CRE (`last_cre_days > 0`) | 1,307 |
| 4. Has CRE within 12 months pre-index | 1,307 |
| 5. ≥ 18 months (548 d) follow-up | 1,307 |
| 6. ≥ 2 post-index CRE measurements | 1,307 |
| 7. Non-missing baseline CRE (90-d window) | 1,307 |
| 8. `eGFR_CRE_baseline ≥ 10` | **1,307** |

## Cohort restriction before matching

| Step | Control n |
|---|---|
| Eligible cohort after `control_group.qmd` build | 1,307 |
| Drop Index year ≤ 2015 (no treated counterpart) | −338 → 969 |
| Drop `has_pre_cre == 0` | (computed in §3) |
| **Eligible for matching** | see chunk `cohort-restrict` |

## Key design choices

| Choice | Reason |
|---|---|
| Drop pre-2016 controls | Treated cohort begins 2016; earlier controls have no temporal counterpart |
| Variable 1:k matching with `ratio=2`, `min.controls=1`, `max.controls=3` | Allows up to 1:3 in the early era where controls are abundant; falls back to 1:1 in the late era (2020–2023) where controls are scarce. Keeps all 379 eligible treated patients in the matched cohort. |
| Caliper 0.2 × SD(logit PS) | Austin (2011) — empirically optimal for bias removal |
| Exact-match on era only | Era shift is large/structural; sex balances adequately via PS alone |
| Include `n_cre_pre_1yr` in PS model | Surveillance intensity is a study-specific candidate confounder; must be balanced explicitly |
| Use binary `white` (1 = White, 0 = other) instead of multi-level `Race` | ≥96% of both cohorts are White; treating Race as a 4-level factor wastes degrees of freedom on near-empty cells. The binary form captures all of the between-group variation that exists. |
| Exclude `Ethnic_Group` from PS model | Highly collinear with `Race`/`white` after harmonisation and contributes little independent variation |
| Exclude `pre_history_days` from PS model | Largely redundant with `n_cre_pre_1yr`; including both inflates the PS model without improving balance |
| Exclude `Cancer_Diagnosis` / `Cancers_stage` | Control cohort lacks these fields (only treated has them) |
| Exclude `pre_median_HGB_90days`, `pre_median_ALB_90days` from PS model | High missingness in controls (routine pre-ICI oncology workup labs are rarely drawn in the control pool); presence would leak treatment status. Shown in Table 1 for transparency only |
| Exclude post-exposure variables (`last_cre_days`, `aki`, outcomes) | Would induce collider bias if included |

## Note on the 2020–2023 era

Within the 2020–2023 stratum the eligible counts are roughly **1:1.4** (treated 142 vs control 192), so a fixed 1:3 match is structurally infeasible — there simply aren't 426 controls available. The variable-ratio specification handles this automatically:

- 2016–2019 era: target up to 1:3 (control supply is plentiful at 777).
- 2020–2023 era: most treated patients will receive **1:1** matches because `min.controls = 1` is the binding constraint once the closest controls are consumed.

This is the right behaviour for this study — it preserves all treated patients, keeps the closest-PS controls in each era, and accepts a smaller k where the data demand it. The downstream Cox / competing-risk model uses subclass-clustered SEs (or `MatchIt` weights), so unequal cluster sizes are handled correctly.

---

::: {.callout-note}
## Prerequisite
Before rendering this page, re-render `treated_group.qmd` and `control_group.qmd` so the new `n_cre_pre_1yr`, `pre_history_days`, `has_pre_cre`, `cre_rate_pre`, `n_cre_post`, `cre_rate_post` columns are present in the output `.xlsx` files.
:::

---

# 1 Setup

```{r}
#| label: setup
suppressPackageStartupMessages({
  library(readxl)
  library(writexl)
  library(dplyr)
  library(tidyr)
  library(MatchIt)
  library(cobalt)
  library(gtsummary)
  library(ggplot2)
})
```

```{r}
#| label: load-data
treated <- read_excel("treated_group.xlsx")
control <- read_excel("control_group.xlsx")

cat("Treated raw n =", nrow(treated), "\n")
cat("Control raw n =", nrow(control), "\n")
```

---

# 2 Harmonize and combine

Both cohorts now carry the binary `white` variable (1 if Race == "White", else 0); no rename is needed for matching.

```{r}
#| label: harmonize
# Columns needed for PS model + bookkeeping + downstream time-to-event
common_vars <- c(
  "EMPI", "Index_date", "last_cre_days", "Age", "male", "white",
  "dm", "htn", "cad", "cirrhosis",
  "ace_arb", "diu", "ppi", "steroids", "smoking", "statins",
  "eGFR_CRE_baseline",
  "pre_median_HGB_90days", "pre_median_ALB_90days",
  "n_cre_pre_1yr", "has_pre_cre"
)

missing_t <- setdiff(common_vars, names(treated))
missing_c <- setdiff(common_vars, names(control))
if (length(missing_t) > 0) stop("Treated missing columns: ", paste(missing_t, collapse = ", "))
if (length(missing_c) > 0) stop("Control missing columns: ", paste(missing_c, collapse = ", "))

combined <- bind_rows(
  treated %>% mutate(treat = 1L) %>% select(all_of(c(common_vars, "treat"))),
  control %>% mutate(treat = 0L) %>% select(all_of(c(common_vars, "treat")))
)

cat("Combined n =", nrow(combined), "\n")
```

---

# 3 Cohort restriction

```{r}
#| label: cohort-restrict
n0 <- nrow(combined)

# Drop controls with Index year <= 2015
combined <- combined %>%
  mutate(index_year = as.numeric(format(as.Date(Index_date), "%Y"))) %>%
  filter(!(treat == 0 & index_year <= 2015))
n1 <- nrow(combined)

# Eligibility: must have pre-index CRE surveillance
combined <- combined %>% filter(has_pre_cre == 1)
n2 <- nrow(combined)

# Era bins
combined <- combined %>%
  mutate(era = factor(ifelse(index_year <= 2019, "2016-2019", "2020-2023")))

# Attrition
attrition <- tibble::tibble(
  Step = c("Combined (treated + control)",
           "Drop controls with Index year <= 2015",
           "Drop has_pre_cre == 0"),
  N = c(n0, n1, n2),
  Dropped = c(NA, n0 - n1, n1 - n2)
)
knitr::kable(attrition)

cat("\nBy group x era:\n")
combined %>% count(treat, era) %>% knitr::kable()
```

---

# 4 Pre-match Table 1

```{r}
#| label: pre-match-table1
pre_tbl <- combined %>%
  select(treat, Age, male, white,
         dm, htn, cad, cirrhosis,
         ace_arb, diu, ppi, steroids, smoking, statins,
         eGFR_CRE_baseline,
         pre_median_HGB_90days, pre_median_ALB_90days,
         n_cre_pre_1yr,
         era) %>%
  tbl_summary(
    by = treat,
    statistic = list(all_continuous() ~ "{median} ({p25}, {p75})"),
    missing_text = "Missing"
  ) %>%
  add_difference() %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels()
pre_tbl
```

---

# 5 Propensity-score model

Drop incomplete cases (PS model + `MatchIt` both require complete covariates).

```{r}
#| label: complete-cases
ps_covars <- c("Age", "male", "white",
               "dm", "htn", "cad", "cirrhosis",
               "ace_arb", "diu", "ppi", "steroids", "smoking", "statins",
               "eGFR_CRE_baseline",
               "n_cre_pre_1yr")
# Excluded from PS model:
# - HGB / ALB: high missingness in controls (kept in Table 1 only).
# - Ethnic_Group: redundant with Race/white after harmonisation.
# - pre_history_days: redundant with n_cre_pre_1yr.

n_before_cc <- nrow(combined)
combined_cc <- combined %>% drop_na(all_of(ps_covars))
n_after_cc <- nrow(combined_cc)

cat("Dropped", n_before_cc - n_after_cc,
    "rows with missing PS covariates.",
    n_after_cc, "remain.\n")

cat("\nMissingness per covariate (full combined):\n")
combined %>% summarise(across(all_of(ps_covars), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "var", values_to = "n_missing") %>%
  arrange(desc(n_missing)) %>%
  knitr::kable()
```

```{r}
#| label: ps-model
ps_formula <- as.formula(paste("treat ~", paste(ps_covars, collapse = " + ")))

ps_fit <- glm(ps_formula, data = combined_cc, family = binomial(link = "logit"))

combined_cc <- combined_cc %>%
  mutate(ps = predict(ps_fit, type = "response"),
         logit_ps = predict(ps_fit, type = "link"))

ggplot(combined_cc, aes(x = logit_ps, fill = factor(treat))) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(values = c("0" = "steelblue", "1" = "firebrick"),
                    labels = c("Control", "Treated"),
                    name = "") +
  labs(x = "Logit(PS)", y = "Density",
       title = "Propensity-score distribution by group") +
  theme_minimal()
```

---

# 6 Matching

Variable-ratio 1:k nearest-neighbor (k between 1 and 3), exact-matched on `era`, caliper 0.2 SD(logit PS).

```{r}
#| label: matchit
set.seed(42)

m.out <- matchit(
  ps_formula,
  data         = combined_cc,
  method       = "nearest",
  distance     = "glm",
  link         = "logit",
  ratio        = 2,
  min.controls = 1,
  max.controls = 3,
  exact        = ~ era,
  caliper      = 0.2,
  std.caliper  = TRUE,
  replace      = FALSE
)

summary(m.out, un = TRUE)
```

```{r}
#| label: matched-data
matched_data <- match.data(m.out)

cat("Matched treated n =", sum(matched_data$treat == 1), "\n")
cat("Matched control n =", sum(matched_data$treat == 0), "\n")
cat("Unmatched treated n =", sum(combined_cc$treat == 1) - sum(matched_data$treat == 1), "\n")
```

---

# 7 Post-match Table 1

```{r}
#| label: post-match-table1
post_tbl <- matched_data %>%
  select(treat, Age, male, white,
         dm, htn, cad, cirrhosis,
         ace_arb, diu, ppi, steroids, smoking, statins,
         eGFR_CRE_baseline,
         pre_median_HGB_90days, pre_median_ALB_90days,
         n_cre_pre_1yr,
         era) %>%
  tbl_summary(
    by = treat,
    statistic = list(all_continuous() ~ "{median} ({p25}, {p75})"),
    missing_text = "Missing"
  ) %>%
  add_difference() %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels()
post_tbl
```

---

# 8 Balance plot

```{r}
#| label: love-plot
#| fig-height: 8
love.plot(
  m.out,
  stats       = c("mean.diffs"),
  threshold   = 0.1,
  abs         = TRUE,
  var.order   = "unadjusted",
  line        = TRUE,
  colors      = c("firebrick", "steelblue"),
  sample.names = c("Unmatched", "Matched")
) +
  labs(title = "Standardized mean differences — pre vs post matching")
```

---

# 9 Export matched cohort

```{r}
#| label: export
# Attach PS, weights, subclass for downstream use
out_df <- matched_data

write_xlsx(out_df, "matched_cohort.xlsx")
cat("Saved matched_cohort.xlsx with", nrow(out_df), "rows and",
    ncol(out_df), "columns.\n")
```

---

# 10 Notes and caveats

- **Unmatched treated**: any treated patient not matched is reported in §6; if >5%, revisit the caliper or widen era bins.
- **Sex not exact-matched**: balanced via PS model alone. Check post-match SMD for `male` in §7; if >0.1, add `exact = ~ era + male`.
- **Cancer type / stage** are only captured for treated patients and so cannot enter the PS model. Residual confounding by cancer type is a study limitation — address in the Discussion.
- **Post-exposure variables** (`last_cre_days`, `aki`, any `*_Xyear` outcome) are excluded from the PS model to avoid collider / post-treatment bias.
- **Use of matched cohort downstream**: Cox / competing-risk models should either (a) use the matched sample with robust/cluster-adjusted SEs by `subclass`, or (b) use weights from `match.data()` in a weighted outcome model.
