---
title: "Results — Outcome Analyses on Matched Cohort"
author: "Tianqi Ouyang"
date: "2026-04-15"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
execute:
  warning: false
  message: false
---

# Summary

Outcome analyses run on the **matched cohort** built in [Matching](matching.qmd) (`matched_cohort.xlsx`).  Pipeline mirrors Kavita's master analysis (Fine-Gray subdistribution + cause-specific Cox + 5-year logistic regression + CIF plots), but uses the propensity-score-matched treated vs control cohort instead of the full ICI-treated cohort.

## Outcomes analyzed

| Outcome | Definition |
|---|---|
| `ckd_composite` | Any of CKD incidence, CKD progression, ESKD composite |
| `ckd_incidence` | >25% eGFR drop AND eGFR <60, sustained ≥90 days, baseline ≥60 |
| `ckd_progression` | >30% eGFR drop, sustained ≥90 days, baseline <60 |
| `eskd_composite` | ESKD by eGFR <10 sustained ≥90 days OR ESRD/KT after index |
| `death` | All-cause mortality (`time_to_death`, `death`) — no competing risk; KM + standard Cox |

## Models per outcome

1. **Cumulative incidence function (CIF) plot** — by `treat` group, competing risk = death.
2. **Univariate Fine-Gray** — `treat` only.
3. **Multivariate Fine-Gray** — `treat` + Age, sex, white, comorbidities, baseline eGFR, surveillance.
4. **Univariate cause-specific Cox** — `treat` only.
5. **Multivariate cause-specific Cox** — same covariates as Fine-Gray multivariate.
6. **5-year logistic regression** — same covariates.

## Primary covariate set

`treat` (1 = ICI-treated, 0 = matched control), `Age`, `male`, `white`, `dm`, `htn`, `cad`, `cirrhosis`, `ppi`, `steroids`, `statins`, `smoking`, `eGFR_CRE_baseline_decrease_10`, `n_cre_pre_1yr`.

Variables excluded relative to Kavita's models: `ICI_Class`, `cancer_group`, individual chemotherapeutic agents, `nephrotoxic_chemo` — these are treated-only and cannot be balanced across groups.

---

# 1 Setup

```{r}
#| label: setup
suppressPackageStartupMessages({
  library(readxl)
  library(dplyr)
  library(tidyr)
  library(survival)
  library(tidycmprsk)
  library(ggsurvfit)
  library(gtsummary)
  library(ggplot2)
  library(broom)
})
```

```{r}
#| label: load
m <- read_excel("matched_cohort.xlsx") %>%
  mutate(
    treat = factor(treat, levels = c(0, 1), labels = c("Control", "ICI Treated")),
    male  = as.integer(male),
    white = as.integer(white),
    across(c(dm, htn, cad, cirrhosis, ppi, steroids, statins, smoking,
             ace_arb, diu),
           ~ as.integer(.)),
    # Convert time variables to years for plotting
    time_to_ckd_cmp_yr = time_to_ckd_cmp / 365.25
  )

cat("Matched cohort: n =", nrow(m), "\n")
cat("By treat:\n"); print(m %>% count(treat))
```

---

# 2 Helper: build outcome-specific competing-risk status

```{r}
#| label: build-status
build_cmp <- function(data, time_var, event_label) {
  # time_var:    e.g. "time_to_ckd_incidence", "time_to_ckd_composite"
  # event_label: matches outcomes[[nm]]$label, e.g. "CKD Composite"
  data %>%
    mutate(
      time_event = pmin(.data[[time_var]], last_cre_days, death_days, na.rm = TRUE),
      status_cmp = case_when(
        !is.na(death_days) & time_event == death_days ~ 2,
        !is.na(.data[[time_var]]) & time_event == .data[[time_var]] ~ 1,
        TRUE ~ 0
      ),
      status_cmp = factor(status_cmp, levels = c(0, 1, 2),
                          labels = c("Censored", event_label, "Death")),
      time_event_yr = time_event / 365.25
    )
}

outcomes <- list(
  ckd_composite   = list(time = "time_to_ckd_composite",   y5 = "ckd_composite_5year",   label = "CKD Composite"),
  ckd_incidence   = list(time = "time_to_ckd_incidence",   y5 = "ckd_incidence_5year",   label = "CKD Incidence"),
  ckd_progression = list(time = "time_to_ckd_progression", y5 = "ckd_progression_5year", label = "CKD Progression"),
  eskd_composite  = list(time = "time_to_eskd_composite",  y5 = "eskd_composite_5year",  label = "ESKD Composite")
)
```

---

# 3 Cumulative incidence functions

```{r}
#| label: cif-plots
#| fig-height: 5
#| fig-width: 7
ps_covars_cif <- c("Age", "male", "white", "dm", "htn", "cad", "cirrhosis",
                   "ppi", "steroids", "statins", "smoking",
                   "eGFR_CRE_baseline_decrease_10", "n_cre_pre_1yr")

for (nm in names(outcomes)) {
  o   <- outcomes[[nm]]
  d   <- build_cmp(m, o$time, o$label)
  cif <- cuminc(Surv(time_event_yr, status_cmp) ~ treat, data = d)

  p <- cif %>%
    ggcuminc(outcome = o$label) +
    add_confidence_interval() +
    add_risktable() +
    scale_x_continuous(breaks = 0:10, limits = c(0, 10)) +
    labs(
      title = paste0("CIF — ", o$label, " by group"),
      x = "Years since Index_date",
      y = paste0("Cumulative incidence of ", o$label)
    ) +
    theme_minimal()

  print(p)
}
```

---

# 4 Fine-Gray subdistribution hazard models

Reports `subdistribution HR (95% CI)` for `treat` (ICI vs Control). Multivariate adjusts for the standard covariate set.

```{r}
#| label: fine-gray
#| results: asis
ps_covars <- c("Age", "male", "white",
               "dm", "htn", "cad", "cirrhosis",
               "ppi", "steroids", "statins", "smoking",
               "eGFR_CRE_baseline_decrease_10", "n_cre_pre_1yr")
mv_rhs <- paste(c("treat", ps_covars), collapse = " + ")

for (nm in names(outcomes)) {
  o <- outcomes[[nm]]
  d <- build_cmp(m, o$time, o$label)
  cat("\n\n## ", o$label, "\n\n")

  # Univariate
  uv_form <- as.formula(paste0("Surv(time_event, status_cmp) ~ treat"))
  uv_fit  <- tidycmprsk::crr(uv_form, data = d,
                             failcode = o$label, cencode = "Censored")
  cat("**Univariate**\n\n")
  print(uv_fit %>% tbl_regression(exp = TRUE) %>% add_n() %>%
          modify_caption(paste0("Fine-Gray — ", o$label, " (univariate)")))

  # Multivariate
  mv_form <- as.formula(paste0("Surv(time_event, status_cmp) ~ ", mv_rhs))
  mv_fit  <- tryCatch(
    tidycmprsk::crr(mv_form, data = d,
                    failcode = o$label, cencode = "Censored"),
    error = function(e) { cat("Multivariate failed:", conditionMessage(e), "\n"); NULL }
  )
  if (!is.null(mv_fit)) {
    cat("\n\n**Multivariate**\n\n")
    print(mv_fit %>% tbl_regression(exp = TRUE) %>% add_n() %>%
            modify_caption(paste0("Fine-Gray — ", o$label, " (multivariate)")))
  }
}
```

---

# 5 Cause-specific Cox models

Reports `cause-specific HR (95% CI)` for the event of interest, treating death as a censoring event.

```{r}
#| label: cause-specific
#| results: asis
for (nm in names(outcomes)) {
  o <- outcomes[[nm]]
  d <- build_cmp(m, o$time, o$label) %>%
    mutate(event_only = as.integer(status_cmp == o$label))
  cat("\n\n## ", o$label, "\n\n")

  # Univariate
  uv_form <- as.formula("Surv(time_event, event_only) ~ treat")
  uv_fit  <- coxph(uv_form, data = d)
  cat("**Univariate**\n\n")
  print(uv_fit %>% tbl_regression(exp = TRUE) %>% add_n() %>%
          modify_caption(paste0("Cause-specific Cox — ", o$label, " (univariate)")))

  # Multivariate
  mv_form <- as.formula(paste0("Surv(time_event, event_only) ~ ", mv_rhs))
  mv_fit  <- tryCatch(coxph(mv_form, data = d),
                      error = function(e) { cat("Failed:", conditionMessage(e), "\n"); NULL })
  if (!is.null(mv_fit)) {
    cat("\n\n**Multivariate**\n\n")
    print(mv_fit %>% tbl_regression(exp = TRUE) %>% add_n() %>%
            modify_caption(paste0("Cause-specific Cox — ", o$label, " (multivariate)")))
  }
}
```

---

# 6 5-year logistic regression

```{r}
#| label: logistic
#| results: asis
for (nm in names(outcomes)) {
  o <- outcomes[[nm]]
  cat("\n\n## ", o$label, " — 5-year incidence\n\n")
  d <- m %>%
    mutate(y5 = as.integer(as.character(.data[[o$y5]])) ) %>%
    filter(!is.na(y5))

  # Univariate
  uv_fit <- glm(y5 ~ treat, data = d, family = binomial)
  cat("**Univariate**\n\n")
  print(uv_fit %>% tbl_regression(exp = TRUE) %>% add_n() %>%
          modify_caption(paste0("Logistic — ", o$label, " 5-year (univariate)")))

  # Multivariate
  mv_form <- as.formula(paste0("y5 ~ ", mv_rhs))
  mv_fit  <- tryCatch(glm(mv_form, data = d, family = binomial),
                      error = function(e) { cat("Failed:", conditionMessage(e), "\n"); NULL })
  if (!is.null(mv_fit)) {
    cat("\n\n**Multivariate**\n\n")
    print(mv_fit %>% tbl_regression(exp = TRUE) %>% add_n() %>%
            modify_caption(paste0("Logistic — ", o$label, " 5-year (multivariate)")))
  }
}
```

---

# 7 Death (all-cause mortality)

Death is analyzed as its own primary endpoint (no competing risk).

## 7.1 Kaplan-Meier survival by `treat`

```{r}
#| label: km-death
#| fig-height: 5
#| fig-width: 7
d_death <- m %>%
  mutate(
    death_event   = as.integer(death == 1),
    time_event    = pmin(time_to_death, last_cre_days, na.rm = TRUE),
    time_event_yr = time_event / 365.25,
    death_5year   = as.integer(time_event <= 5 * 365.25 & death_event == 1)
  )

km_fit <- survfit2(Surv(time_event_yr, death_event) ~ treat, data = d_death)

km_fit %>%
  ggsurvfit() +
  add_confidence_interval() +
  add_risktable() +
  scale_x_continuous(breaks = 0:10, limits = c(0, 10)) +
  labs(
    title = "Kaplan-Meier — All-cause mortality by group",
    x = "Years since Index_date",
    y = "Survival probability"
  ) +
  theme_minimal()
```

## 7.2 Cox proportional-hazards models

```{r}
#| label: cox-death
#| results: asis
# Univariate
uv_fit <- coxph(Surv(time_event, death_event) ~ treat, data = d_death)
cat("**Univariate**\n\n")
print(uv_fit %>% tbl_regression(exp = TRUE) %>% add_n() %>%
        modify_caption("Cox — Death (univariate)"))

# Multivariate
mv_form <- as.formula(paste0("Surv(time_event, death_event) ~ ", mv_rhs))
mv_fit  <- tryCatch(coxph(mv_form, data = d_death),
                    error = function(e) { cat("Failed:", conditionMessage(e), "\n"); NULL })
if (!is.null(mv_fit)) {
  cat("\n\n**Multivariate**\n\n")
  print(mv_fit %>% tbl_regression(exp = TRUE) %>% add_n() %>%
          modify_caption("Cox — Death (multivariate)"))
}
```

## 7.3 5-year logistic regression — death within 5 years

```{r}
#| label: logistic-death
#| results: asis
d5 <- d_death %>% filter(!is.na(death_5year))

uv_fit <- glm(death_5year ~ treat, data = d5, family = binomial)
cat("**Univariate**\n\n")
print(uv_fit %>% tbl_regression(exp = TRUE) %>% add_n() %>%
        modify_caption("Logistic — Death 5-year (univariate)"))

mv_form <- as.formula(paste0("death_5year ~ ", mv_rhs))
mv_fit  <- tryCatch(glm(mv_form, data = d5, family = binomial),
                    error = function(e) { cat("Failed:", conditionMessage(e), "\n"); NULL })
if (!is.null(mv_fit)) {
  cat("\n\n**Multivariate**\n\n")
  print(mv_fit %>% tbl_regression(exp = TRUE) %>% add_n() %>%
          modify_caption("Logistic — Death 5-year (multivariate)"))
}
```

---

# 8 Notes

- All models adjust for the propensity-score matched cohort on top of the explicit covariates, providing **doubly-robust** estimates.
- Standard errors do **not** account for matched-pair clustering; for inference critical to publication, refit with `cluster(subclass)` or use the `MatchIt` weights with `survival::coxph(weights = ...)`.
- Treated-only variables (cancer type/stage, ICI class, individual chemotherapy agents, `nephrotoxic_chemo`) cannot be included because they have no analogue in the control group; residual confounding by cancer-type is a study limitation.
- Death is treated as a competing risk in CIF and Fine-Gray models, and as a censoring event in the cause-specific Cox.
- 5-year logistic regression complements the time-to-event models with a fixed-horizon interpretation; patients censored before 5y with no event are dropped.
