---
title: "Results - Outcome Analyses on Matched Cohort"
author: "Tianqi Ouyang"
date: "2026-04-15"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
execute:
  warning: false
  message: false
---

# Summary

Outcome analyses run on the **matched cohort** built in [Matching](matching.qmd) (`matched_cohort.xlsx`).  Pipeline mirrors Kavita's master analysis (Fine-Gray subdistribution + cause-specific Cox + 5-year logistic regression + CIF plots), but uses the propensity-score-matched treated vs control cohort instead of the full ICI-treated cohort.

## Outcomes analyzed

| Outcome | Definition |
|---|---|
| `ckd_composite` | Any of CKD incidence, CKD progression, ESKD composite |
| `ckd_incidence` | >25% eGFR drop AND eGFR <60, sustained >=90 days, baseline >=60 |
| `ckd_progression` | >30% eGFR drop, sustained >=90 days, baseline <60 |
| `eskd_composite` | ESKD by eGFR <10 sustained >=90 days OR ESRD/KT after index |
| `death` | All-cause mortality (`time_to_death`, `death`) - no competing risk; KM + standard Cox |

## Models per outcome

1. **Cumulative incidence function (CIF) plot** - by `treat` group, competing risk = death.
2. **Univariate Fine-Gray** - `treat` only.
3. **Multivariate Fine-Gray** - `treat` + Age, sex, white, comorbidities, baseline eGFR, surveillance.
4. **Univariate cause-specific Cox** - `treat` only.
5. **Multivariate cause-specific Cox** - same covariates as Fine-Gray multivariate.
6. **5-year logistic regression** - same covariates.

## Primary covariate set

`treat` (1 = ICI-treated, 0 = matched control), `Age`, `male`, `white`, `dm`, `htn`, `cad`, `cirrhosis`, `ppi`, `steroids`, `statins`, `smoking`, `eGFR_CRE_baseline_decrease_10`, `n_cre_pre_1yr`.

Variables excluded relative to Kavita's models: `ICI_Class`, `cancer_group`, individual chemotherapeutic agents, `nephrotoxic_chemo` - these are treated-only and cannot be balanced across groups.

---

# Statistical Methods (manuscript-ready)

## Study population

Adult patients who received immune-checkpoint-inhibitor (ICI) therapy at Mass General Brigham between 2016 and 2023 (n = 379) were compared with a contemporaneous non-ICI control cohort of patients with a primary neoplasm diagnosis (ICD-10 C00-D49) identified from the Research Patient Data Registry (RPDR). The index date was defined as the first ICI dose for the treated group and the first qualifying neoplasm-diagnosis date for the control group. Cohort inclusion criteria for both groups required (i) a baseline serum creatinine within 12 months prior to the index date, (ii) at least two post-index creatinine measurements, (iii) >= 18 months of follow-up, and (iv) a baseline estimated glomerular filtration rate (eGFR) >= 10 mL/min/1.73 m2. Controls with an index date before 2016 (n = 338) were excluded a priori because they had no temporal counterpart in the treated cohort. After all inclusion and exclusion criteria, 379 treated and 969 control patients were eligible for matching.

## Propensity-score matching

Treatment-effect estimation used a propensity-score (PS) matched cohort to mitigate baseline imbalance. The PS for receiving ICI therapy was estimated using multivariable logistic regression with the following pre-index covariates: age, male sex, white race (binary), diabetes, hypertension, coronary artery disease, cirrhosis, baseline use of proton-pump inhibitors, systemic corticosteroids, statins, and active smoking; the baseline CKD-EPI 2021 eGFR (creatinine-only, race-free); and the number of pre-index creatinine measurements within 365 days (`n_cre_pre_1yr`) as an explicit adjustment for differential ascertainment intensity. Variable-ratio 1:k nearest-neighbor matching was performed with `MatchIt` (R) using the logit of the propensity score, no replacement, a caliper of 0.2 x SD(logit PS) (Austin 2011), exact matching on calendar era (2016-2019 vs 2020-2023), and `ratio = 2`, `min.controls = 1`, `max.controls = 3` to maintain feasibility in the 2020-2023 stratum where the control pool was structurally limited. Balance was assessed by standardized mean differences (SMDs), with a target of |SMD| < 0.1 for all matching variables.

## Outcomes

Five outcomes were analyzed: (1) CKD composite (CKD incidence, CKD progression, or ESKD composite); (2) CKD incidence (>25% relative eGFR drop **and** eGFR < 60 mL/min/1.73 m2 sustained >= 90 days, in patients with baseline eGFR >= 60); (3) CKD progression (>30% relative eGFR drop sustained >= 90 days in patients with baseline eGFR < 60); (4) ESKD composite (eGFR < 10 sustained >= 90 days **or** documented end-stage renal disease / kidney transplant after the index date); and (5) all-cause mortality. Time-to-event was measured in days from the index date, with administrative censoring at the last available creatinine date (`last_cre_days`). All eGFR values were computed with the 2021 CKD-EPI creatinine-only equation (Inker et al. 2021), which is race-free.

## Time-to-event modelling of CKD/ESKD outcomes

For each CKD/ESKD endpoint, death was treated as a competing risk. Cumulative incidence functions (CIFs) were estimated non-parametrically using the Aalen-Johansen estimator (`tidycmprsk::cuminc`) and plotted with 95% point-wise confidence intervals. Subdistribution hazard ratios were estimated with Fine-Gray models (`tidycmprsk::crr`) in unadjusted and adjusted forms. The adjusted model included treatment group plus age, male sex, white race, diabetes, hypertension, coronary artery disease, cirrhosis, baseline PPI / steroid / statin use, smoking, baseline eGFR (expressed per 10-mL decrement: `eGFR_CRE_baseline_decrease_10`), and `n_cre_pre_1yr`. Cause-specific hazard ratios were estimated with Cox proportional-hazards models (`survival::coxph`), censoring death; the same covariate set was used. The proportional-hazards assumption was inspected by Schoenfeld residuals.

## All-cause mortality

Survival was summarised by Kaplan-Meier estimation (`survfit2`). Hazard ratios for mortality were estimated with Cox proportional-hazards models, treating death as the primary event without competing risk.

## Fixed-horizon (5-year) analyses

In parallel with the time-to-event models, 5-year event probabilities were modelled by logistic regression for each outcome (univariate and multivariate using the same covariate set). Patients censored before 5 years without the event were excluded from the 5-year denominator.

## Longitudinal eGFR slope

To quantify annual eGFR decline between groups, every post-index serum-creatinine measurement was extracted from the RPDR Lab files and converted to an eGFR with CKD-EPI 2021, advancing chronologic age to the lab date (treated: 9,455 observations; control: 11,148 observations). Three linear mixed-effects models (`lme4::lmer`, REML, `bobyqa` optimizer; Satterthwaite degrees of freedom via `lmerTest`) were fit:

- **Long-term**: all post-index observations, `eGFR ~ time x treat + (time | EMPI)` - random intercept and slope per patient.
- **First-year**: observations within 365 days of index, `eGFR ~ time x treat + (1 | EMPI)` - random intercept only (slope not identifiable because 55 % of controls had only one first-year measurement).
- **Three-year**: observations within 1095 days, `eGFR ~ time x treat + (time | EMPI)` - random intercept and slope.

In each model the `time x treat` interaction term is the between-group difference in annual eGFR decline (mL/min/1.73 m2 per year). Each model's group-mean trajectory is visualised with the fixed-effects marginal prediction (and 95% CI ribbon) from the fitted `lmer` - no LOESS smoother, no per-patient slope plot.

## Software and inference

All analyses were performed in R using `MatchIt`, `survival`, `tidycmprsk`, `lme4`, `lmerTest`, `ggsurvfit`, `gtsummary`, `cobalt`, and `tidyverse`. Two-sided p-values < 0.05 were considered statistically significant. Standard errors in the matched-cohort outcome models do **not** explicitly account for the matched-pair clustering induced by 1:k matching; sensitivity analyses with subclass-clustered robust standard errors (`coxph(... , cluster = subclass)`) or `MatchIt` matching weights are planned for the final manuscript. Treated-only covariates (cancer type and stage, ICI class, individual chemotherapeutic agents, nephrotoxic chemotherapy flag) have no analogue in the control group and were therefore excluded from the PS model and outcome adjustment; this is acknowledged as a source of residual confounding by indication.

---

# 1 Setup

```{r}
#| label: setup
suppressPackageStartupMessages({
  library(readxl)
  library(dplyr)
  library(tidyr)
  library(survival)
  library(tidycmprsk)
  library(ggsurvfit)
  library(gtsummary)
  library(ggplot2)
  library(broom)
})
```

```{r}
#| label: load
m <- read_excel("matched_cohort.xlsx") %>%
  mutate(
    treat = factor(treat, levels = c(0, 1), labels = c("Control", "ICI Treated")),
    male  = as.integer(male),
    white = as.integer(white),
    across(c(dm, htn, cad, cirrhosis, ppi, steroids, statins, smoking,
             ace_arb, diu),
           ~ as.integer(.)),
    # Convert time variables to years for plotting
    time_to_ckd_cmp_yr = time_to_ckd_cmp / 365.25
  )

cat("Matched cohort: n =", nrow(m), "\n")
cat("By treat:\n"); print(m %>% count(treat))
```

---

# 2 Helper: build outcome-specific competing-risk status

```{r}
#| label: build-status
build_cmp <- function(data, time_var, event_label) {
  # time_var:    e.g. "time_to_ckd_incidence", "time_to_ckd_composite"
  # event_label: matches outcomes[[nm]]$label, e.g. "CKD Composite"
  data %>%
    mutate(
      time_event = pmin(.data[[time_var]], last_cre_days, death_days, na.rm = TRUE),
      status_cmp = case_when(
        !is.na(death_days) & time_event == death_days ~ 2,
        !is.na(.data[[time_var]]) & time_event == .data[[time_var]] ~ 1,
        TRUE ~ 0
      ),
      status_cmp = factor(status_cmp, levels = c(0, 1, 2),
                          labels = c("Censored", event_label, "Death")),
      time_event_yr = time_event / 365.25
    )
}

outcomes <- list(
  ckd_composite   = list(time = "time_to_ckd_composite",   y5 = "ckd_composite_5year",   label = "CKD Composite"),
  ckd_incidence   = list(time = "time_to_ckd_incidence",   y5 = "ckd_incidence_5year",   label = "CKD Incidence"),
  ckd_progression = list(time = "time_to_ckd_progression", y5 = "ckd_progression_5year", label = "CKD Progression"),
  eskd_composite  = list(time = "time_to_eskd_composite",  y5 = "eskd_composite_5year",  label = "ESKD Composite")
)
```

---

# 3 Cumulative incidence functions

```{r}
#| label: cif-plots
#| fig-height: 5
#| fig-width: 7
ps_covars_cif <- c("Age", "male", "white", "dm", "htn", "cad", "cirrhosis",
                   "ppi", "steroids", "statins", "smoking",
                   "eGFR_CRE_baseline_decrease_10", "n_cre_pre_1yr")

for (nm in names(outcomes)) {
  o   <- outcomes[[nm]]
  d   <- build_cmp(m, o$time, o$label)
  cif <- cuminc(Surv(time_event_yr, status_cmp) ~ treat, data = d)

  p <- cif %>%
    ggcuminc(outcome = o$label) +
    add_confidence_interval() +
    add_risktable() +
    scale_x_continuous(breaks = 0:10, limits = c(0, 10)) +
    labs(
      title = paste0("CIF: ", o$label, " by group"),
      x = "Years since Index_date",
      y = paste0("Cumulative incidence of ", o$label)
    ) +
    theme_minimal()

  print(p)
}
```

---

# 4 Fine-Gray subdistribution hazard models

Reports subdistribution HR (95% CI) for `treat` (ICI vs Control). Univariate (left) and multivariate (right) shown side-by-side; rows grouped by outcome. Multivariate adjusts for the standard covariate set.

```{r}
#| label: fine-gray-build
ps_covars <- c("Age", "male", "white",
               "dm", "htn", "cad", "cirrhosis",
               "ppi", "steroids", "statins", "smoking",
               "eGFR_CRE_baseline_decrease_10", "n_cre_pre_1yr")
mv_rhs <- paste(c("treat", ps_covars), collapse = " + ")

fg_uv_tbls <- lapply(outcomes, function(o) {
  d <- build_cmp(m, o$time, o$label)
  tidycmprsk::crr(
    Surv(time_event, status_cmp) ~ treat, data = d,
    failcode = o$label, cencode = "Censored"
  ) %>% tbl_regression(exp = TRUE) %>% add_n()
})

fg_mv_tbls <- lapply(outcomes, function(o) {
  d <- build_cmp(m, o$time, o$label)
  tryCatch(
    tidycmprsk::crr(
      as.formula(paste0("Surv(time_event, status_cmp) ~ ", mv_rhs)),
      data = d, failcode = o$label, cencode = "Censored"
    ) %>% tbl_regression(exp = TRUE) %>% add_n(),
    error = function(e) NULL
  )
})
```

### Univariate

```{r}
#| label: fg-uv
tbl_stack(fg_uv_tbls, group_header = sapply(outcomes, `[[`, "label")) %>%
  modify_caption("Fine-Gray subdistribution HR - univariate (treat only)")
```

### Multivariate

```{r}
#| label: fg-mv
tbl_stack(fg_mv_tbls[!sapply(fg_mv_tbls, is.null)],
          group_header = sapply(outcomes[!sapply(fg_mv_tbls, is.null)], `[[`, "label")) %>%
  modify_caption("Fine-Gray subdistribution HR - multivariate (adjusted)")
```

---

# 5 Cause-specific Cox models

Reports cause-specific HR (95% CI) for the event of interest, treating death as a censoring event.

```{r}
#| label: cause-specific-build
cs_uv_tbls <- lapply(outcomes, function(o) {
  d <- build_cmp(m, o$time, o$label) %>%
    mutate(event_only = as.integer(status_cmp == o$label))
  coxph(Surv(time_event, event_only) ~ treat, data = d) %>%
    tbl_regression(exp = TRUE) %>% add_n()
})

cs_mv_tbls <- lapply(outcomes, function(o) {
  d <- build_cmp(m, o$time, o$label) %>%
    mutate(event_only = as.integer(status_cmp == o$label))
  tryCatch(
    coxph(as.formula(paste0("Surv(time_event, event_only) ~ ", mv_rhs)), data = d) %>%
      tbl_regression(exp = TRUE) %>% add_n(),
    error = function(e) NULL
  )
})
```

### Univariate

```{r}
#| label: cs-uv
tbl_stack(cs_uv_tbls, group_header = sapply(outcomes, `[[`, "label")) %>%
  modify_caption("Cause-specific HR - univariate (treat only)")
```

### Multivariate

```{r}
#| label: cs-mv
tbl_stack(cs_mv_tbls[!sapply(cs_mv_tbls, is.null)],
          group_header = sapply(outcomes[!sapply(cs_mv_tbls, is.null)], `[[`, "label")) %>%
  modify_caption("Cause-specific HR - multivariate (adjusted)")
```

---

# 6 5-year logistic regression

```{r}
#| label: logistic-build
lg_uv_tbls <- lapply(outcomes, function(o) {
  d <- m %>%
    mutate(y5 = as.integer(as.character(.data[[o$y5]]))) %>%
    filter(!is.na(y5))
  glm(y5 ~ treat, data = d, family = binomial) %>%
    tbl_regression(exp = TRUE) %>% add_n()
})

lg_mv_tbls <- lapply(outcomes, function(o) {
  d <- m %>%
    mutate(y5 = as.integer(as.character(.data[[o$y5]]))) %>%
    filter(!is.na(y5))
  tryCatch(
    glm(as.formula(paste0("y5 ~ ", mv_rhs)), data = d, family = binomial) %>%
      tbl_regression(exp = TRUE) %>% add_n(),
    error = function(e) NULL
  )
})
```

### Univariate

```{r}
#| label: lg-uv
tbl_stack(lg_uv_tbls, group_header = sapply(outcomes, `[[`, "label")) %>%
  modify_caption("5-year incidence - univariate logistic OR (treat only)")
```

### Multivariate

```{r}
#| label: lg-mv
tbl_stack(lg_mv_tbls[!sapply(lg_mv_tbls, is.null)],
          group_header = sapply(outcomes[!sapply(lg_mv_tbls, is.null)], `[[`, "label")) %>%
  modify_caption("5-year incidence - multivariate logistic OR (adjusted)")
```

---

# 7 Death (all-cause mortality)

Death is analyzed as its own primary endpoint (no competing risk).

## 7.1 Kaplan-Meier survival by `treat`

```{r}
#| label: km-death
#| fig-height: 5
#| fig-width: 7
d_death <- m %>%
  mutate(
    death_event   = as.integer(death == 1),
    time_event    = pmin(time_to_death, last_cre_days, na.rm = TRUE),
    time_event_yr = time_event / 365.25,
    death_5year   = as.integer(time_event <= 5 * 365.25 & death_event == 1)
  )

km_fit <- survfit2(Surv(time_event_yr, death_event) ~ treat, data = d_death)

km_fit %>%
  ggsurvfit() +
  add_confidence_interval() +
  add_risktable() +
  scale_x_continuous(breaks = 0:10, limits = c(0, 10)) +
  labs(
    title = "Kaplan-Meier: All-cause mortality by group",
    x = "Years since Index_date",
    y = "Survival probability"
  ) +
  theme_minimal()
```

## 7.2 Cox proportional-hazards models

### Univariate

```{r}
#| label: cox-death-uv
coxph(Surv(time_event, death_event) ~ treat, data = d_death) %>%
  tbl_regression(exp = TRUE) %>% add_n() %>%
  modify_caption("Cox HR - Death (univariate)")
```

### Multivariate

```{r}
#| label: cox-death-mv
coxph(as.formula(paste0("Surv(time_event, death_event) ~ ", mv_rhs)),
      data = d_death) %>%
  tbl_regression(exp = TRUE) %>% add_n() %>%
  modify_caption("Cox HR - Death (multivariate, adjusted)")
```

## 7.3 5-year logistic regression - death within 5 years

```{r}
#| label: logistic-death-prep
d5 <- d_death %>% filter(!is.na(death_5year))
```

### Univariate

```{r}
#| label: lg-death-uv
glm(death_5year ~ treat, data = d5, family = binomial) %>%
  tbl_regression(exp = TRUE) %>% add_n() %>%
  modify_caption("Death 5-year - univariate logistic OR")
```

### Multivariate

```{r}
#| label: lg-death-mv
glm(as.formula(paste0("death_5year ~ ", mv_rhs)),
    data = d5, family = binomial) %>%
  tbl_regression(exp = TRUE) %>% add_n() %>%
  modify_caption("Death 5-year - multivariate logistic OR (adjusted)")
```

---

# 8 eGFR slope analysis

Annual eGFR decline (mL/min/year) compared between ICI-treated and matched-control patients, using per-lab CRE observations from the RPDR Lab files (not yearly aggregates). Two models are fit:

| Model | Time window | Random effects |
|---|---|---|
| **8.2** Long-term slope | All post-index CRE labs | Intercept + slope per patient |
| **8.3** First-year slope | Days 1-365 post-index only | Intercept per patient (random slope not estimable for ~55% of controls with only one first-year obs) |

Fixed effects: `eGFR_CRE ~ year * treat`. The `year:treatICI Treated` interaction term **is** the difference in annual GFR decline between groups.

## 8.1 Build per-lab eGFR dataframe

```{r}
#| label: build-lab-egfr
suppressPackageStartupMessages({
  library(lme4)
  library(lmerTest)
})

# CKD-EPI 2021 creatinine-only (race-free)
ckd_epi_gfr <- function(creat, male, age) {
  k <- ifelse(male == 1, 0.9, 0.7)
  a <- ifelse(male == 1, -0.302, -0.241)
  m <- ifelse(male == 1, 1, 1.012)
  142 * pmin(creat / k, 1)^a * pmax(creat / k, 1)^-1.2 * 0.9938^age * m
}

# Treated RPDR Lab
t_lab <- read.table(
  "/Users/to909/Partners HealthCare Dropbox/Tianqi Ouyang/CKD after ICI/RPDR/mes98_093025104659807332_Lab.txt",
  sep = "|", fill = TRUE, quote = "", header = TRUE, stringsAsFactors = FALSE
) %>%
  select(EMPI, Seq_Date_Time, Result, Group_Id) %>%
  filter(Group_Id == "CRE", grepl("^\\d+\\.\\d+$", Result)) %>%
  mutate(
    Seq_Date_Time = as.Date(Seq_Date_Time, "%m/%d/%Y"),
    Result        = as.numeric(Result)
  )

# Control RPDR Lab (04092026 pull)
c_lab <- read.table(
  "/Users/to909/Partners HealthCare Dropbox/Tianqi Ouyang/CKD after ICI/RPDR/RPDR Joyita/04092026/mes98_040726114217167239_Lab.txt",
  sep = "|", fill = TRUE, quote = "", header = TRUE, stringsAsFactors = FALSE
) %>%
  select(EMPI, Seq_Date_Time, Result, Group_Id) %>%
  filter(Group_Id == "CRE", grepl("^\\d+\\.\\d+$", Result)) %>%
  mutate(
    Seq_Date_Time = as.Date(Seq_Date_Time, "%m/%d/%Y"),
    Result        = as.numeric(Result)
  )

# Per-patient anchor (Age, male, Index_date, treat) - one row per EMPI
anchor <- m %>%
  distinct(EMPI, .keep_all = TRUE) %>%
  mutate(Index_date = as.Date(Index_date)) %>%
  select(EMPI, treat, male, Age, Index_date, last_cre_days)

# Join labs to anchor based on treat
labs <- bind_rows(
  t_lab %>% inner_join(anchor %>% filter(treat == "ICI Treated"), by = "EMPI"),
  c_lab %>% inner_join(anchor %>% filter(treat == "Control"),     by = "EMPI")
) %>%
  filter(Seq_Date_Time > Index_date) %>%
  mutate(
    days_since_index = as.numeric(Seq_Date_Time - Index_date),
    year_since_index = days_since_index / 365.25,
    eGFR_CRE         = ckd_epi_gfr(Result, male, Age + year_since_index)
  ) %>%
  filter(eGFR_CRE >= 0, eGFR_CRE <= 200)

cat("Per-lab eGFR observations:\n")
print(labs %>% count(treat))
cat("\nObs per patient (median, IQR):\n")
print(labs %>% group_by(treat, EMPI) %>% summarise(n = n(), .groups = "drop") %>%
        group_by(treat) %>% summarise(
          median_n = median(n), p25 = quantile(n, .25), p75 = quantile(n, .75),
          .groups = "drop"
        ))
```

```{r}
#| label: helper-lmer-pred
# Marginal (fixed-effects-only) prediction with 95% CI from an lmer fit.
# Returns a data.frame: year_since_index, treat, fit, lo, hi.
lmer_marginal_pred <- function(fit, year_grid) {
  grid <- expand.grid(
    year_since_index = year_grid,
    treat            = c("Control", "ICI Treated"),
    stringsAsFactors = FALSE
  )
  grid$treat <- factor(grid$treat, levels = c("Control", "ICI Treated"))

  mm <- model.matrix(~ year_since_index * treat, data = grid)
  fe <- lme4::fixef(fit)
  grid$fit <- as.numeric(mm %*% fe)
  v  <- mm %*% as.matrix(vcov(fit)) %*% t(mm)
  grid$se <- sqrt(diag(v))
  grid$lo <- grid$fit - 1.96 * grid$se
  grid$hi <- grid$fit + 1.96 * grid$se
  grid
}
```

## 8.2 Long-term slope - all post-index data

```{r}
#| label: lmm-all
fit_all <- lmer(
  eGFR_CRE ~ year_since_index * treat + (year_since_index | EMPI),
  data = labs,
  REML = TRUE,
  control = lmerControl(optimizer = "bobyqa")
)

fit_all %>%
  tbl_regression(
    intercept = TRUE,
    label = list(
      year_since_index ~ "Time (yr) - control slope",
      `year_since_index:treat` ~ "Year x ICI Treated (diff in slope)"
    )
  ) %>%
  add_n() %>%
  modify_caption("Linear mixed-effects model - eGFR (mL/min/1.73 m2) over all post-index follow-up")
```

```{r}
#| label: lmer-plot-all
#| fig-height: 5
#| fig-width: 8
pred_all <- lmer_marginal_pred(fit_all, year_grid = seq(0, max(labs$year_since_index, na.rm = TRUE), length.out = 200))

ggplot(pred_all, aes(x = year_since_index, y = fit, color = treat, fill = treat)) +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = .2, color = NA) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = c("Control" = "steelblue", "ICI Treated" = "firebrick")) +
  scale_fill_manual( values = c("Control" = "steelblue", "ICI Treated" = "firebrick")) +
  scale_x_continuous(breaks = 0:10) +
  labs(
    title = "Model-predicted mean eGFR over time by group (long-term lmer)",
    subtitle = "Fixed-effects marginal prediction with 95% CI",
    x = "Years since Index_date",
    y = "Predicted eGFR (mL/min/1.73 m2)",
    color = NULL, fill = NULL
  ) +
  theme_minimal()
```

## 8.3 First-year slope (days 1-365)

```{r}
#| label: lmm-1yr
labs_yr1 <- labs %>% filter(days_since_index <= 365)

fit_yr1 <- lmer(
  eGFR_CRE ~ year_since_index * treat + (1 | EMPI),
  data = labs_yr1,
  REML = TRUE,
  control = lmerControl(optimizer = "bobyqa")
)

fit_yr1 %>%
  tbl_regression(
    intercept = TRUE,
    label = list(
      year_since_index ~ "Time (yr) - control slope",
      `year_since_index:treat` ~ "Year x ICI Treated (diff in slope)"
    )
  ) %>%
  add_n() %>%
  modify_caption("Linear mixed-effects model - first-year (<=365 d) eGFR decline")
```

```{r}
#| label: lmer-plot-yr1
#| fig-height: 5
#| fig-width: 8
pred_yr1 <- lmer_marginal_pred(fit_yr1, year_grid = seq(0, 1, length.out = 200))

ggplot(pred_yr1, aes(x = year_since_index, y = fit, color = treat, fill = treat)) +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = .2, color = NA) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = c("Control" = "steelblue", "ICI Treated" = "firebrick")) +
  scale_fill_manual( values = c("Control" = "steelblue", "ICI Treated" = "firebrick")) +
  scale_x_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1),
                     labels = c("0", "3 mo", "6 mo", "9 mo", "1 yr")) +
  labs(
    title = "Model-predicted mean eGFR over first year (lmer)",
    subtitle = "Fixed-effects marginal prediction with 95% CI",
    x = "Time since Index_date",
    y = "Predicted eGFR (mL/min/1.73 m2)",
    color = NULL, fill = NULL
  ) +
  theme_minimal()
```

## 8.4 Three-year slope (days 1-1095)

```{r}
#| label: lmm-3yr
labs_yr3 <- labs %>% filter(days_since_index <= 1095)

fit_yr3 <- lmer(
  eGFR_CRE ~ year_since_index * treat + (year_since_index | EMPI),
  data = labs_yr3,
  REML = TRUE,
  control = lmerControl(optimizer = "bobyqa")
)

fit_yr3 %>%
  tbl_regression(
    intercept = TRUE,
    label = list(
      year_since_index ~ "Time (yr) - control slope",
      `year_since_index:treat` ~ "Year x ICI Treated (diff in slope)"
    )
  ) %>%
  add_n() %>%
  modify_caption("Linear mixed-effects model - first 3 years (<=1095 d) eGFR decline")
```

```{r}
#| label: lmer-plot-yr3
#| fig-height: 5
#| fig-width: 8
pred_yr3 <- lmer_marginal_pred(fit_yr3, year_grid = seq(0, 3, length.out = 200))

ggplot(pred_yr3, aes(x = year_since_index, y = fit, color = treat, fill = treat)) +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = .2, color = NA) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = c("Control" = "steelblue", "ICI Treated" = "firebrick")) +
  scale_fill_manual( values = c("Control" = "steelblue", "ICI Treated" = "firebrick")) +
  scale_x_continuous(breaks = c(0, 0.5, 1, 1.5, 2, 2.5, 3)) +
  labs(
    title = "Model-predicted mean eGFR over 3 years (lmer)",
    subtitle = "Fixed-effects marginal prediction with 95% CI",
    x = "Years since Index_date",
    y = "Predicted eGFR (mL/min/1.73 m2)",
    color = NULL, fill = NULL
  ) +
  theme_minimal()
```

---

# 9 Notes

- All models adjust for the propensity-score matched cohort on top of the explicit covariates, providing **doubly-robust** estimates.
- Standard errors do **not** account for matched-pair clustering; for inference critical to publication, refit with `cluster(subclass)` or use the `MatchIt` weights with `survival::coxph(weights = ...)`.
- Treated-only variables (cancer type/stage, ICI class, individual chemotherapy agents, `nephrotoxic_chemo`) cannot be included because they have no analogue in the control group; residual confounding by cancer-type is a study limitation.
- Death is treated as a competing risk in CIF and Fine-Gray models, and as a censoring event in the cause-specific Cox.
- 5-year logistic regression complements the time-to-event models with a fixed-horizon interpretation; patients censored before 5y with no event are dropped.
