---
title: "Treated Group – Descriptive Analysis"
author: "Tianqi Ouyang"
date: "2026-03-20"
format:
  html:
    theme: lumen
    toc: true
    toc-float: true
    toc-depth: 3
execute:
  echo: false
  include: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = FALSE, include = FALSE)
```

```{r}
#| label: libraries
rm(list = ls())
library(readxl)
library(dplyr)
library(tidyverse)
library(writexl)
library(gtsummary)
library(survival)
library(ggplot2)
library(survminer)
library(ggsurvfit)
library(tidycmprsk)
library(table1)
library(readr)
```

---

::: {style="display:none;"}

# 2 Functions

Reused from `analysis_ici_ckd.qmd`.

## 2.1 pre_lab_median

```{r}
#| label: fn-pre-lab-median
pre_lab_median <- function(data, month, time, index_date, Group_Id) {
  index_date <- enquo(index_date)
  if (month) {
    time_value <- time * 30
  } else {
    time_value <- time
  }

  median_name <- paste0("pre_median_", Group_Id, "_", time_value, "days")

  result <- data %>%
    filter(Group_Id == !!Group_Id, grepl("^\\d+\\.\\d+$", Result)) %>%
    filter(!is.na(Result)) %>%
    mutate(Result = as.numeric(Result)) %>%
    filter(Seq_Date_Time >= (!!index_date) - time_value,
           Seq_Date_Time <= !!index_date) %>%
    group_by(EMPI) %>%
    summarise(
      !!median_name := median(Result, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    select(EMPI, !!median_name)

  return(result)
}
```

## 2.2 pre_lab_median_year

```{r}
#| label: fn-pre-lab-median-year
pre_lab_median_year <- function(data, year, days, index_date, Group_Id) {
  index_date <- enquo(index_date)

  last_offset  <- year * 365
  first_offset <- last_offset - days

  median_name      <- paste0("pre_median_", Group_Id, "_", year, "year")
  lab_num_name     <- paste0("lab_num_", year, "year")
  closest_name     <- paste0("closest_", Group_Id, "_", year, "year")
  closest_date_name <- paste0("closest_", Group_Id, "_", year, "year_date")

  median <- data %>%
    filter(Group_Id == !!Group_Id, grepl("^\\d+\\.\\d+$", Result)) %>%
    filter(!is.na(Result)) %>%
    mutate(Result = as.numeric(Result)) %>%
    filter(Seq_Date_Time >= (!!index_date + first_offset),
           Seq_Date_Time <= !!index_date + last_offset) %>%
    group_by(EMPI) %>%
    summarise(
      !!median_name := median(Result, na.rm = TRUE),
      !!lab_num_name := n(),
      .groups = "drop"
    )

  closest <- data %>%
    filter(Group_Id == !!Group_Id, grepl("^\\d+\\.\\d+$", Result)) %>%
    filter(!is.na(Result)) %>%
    mutate(Result = as.numeric(Result)) %>%
    filter(Seq_Date_Time <= !!index_date + last_offset) %>%
    arrange(EMPI, desc(Seq_Date_Time)) %>%
    distinct(EMPI, .keep_all = TRUE) %>%
    rename(!!closest_name := Result,
           !!closest_date_name := Seq_Date_Time) %>%
    select(EMPI, !!index_date, !!closest_name, !!closest_date_name)

  result <- median %>%
    merge(closest, by = "EMPI", all = TRUE) %>%
    mutate(
      !!median_name := ifelse(is.na(!!sym(median_name)),
                              !!sym(closest_name),
                              !!sym(median_name))
    ) %>%
    select(EMPI, !!median_name, !!lab_num_name)

  return(result)
}
```

## 2.3 left_join_all

```{r}
#| label: fn-left-join-all
left_join_all <- function(master, df_list, by) {
  for (df in df_list) {
    master <- merge(master, df, by = by, all.x = TRUE)
  }
  return(master)
}
```

## 2.4 get_ckd_stage

```{r}
#| label: fn-get-ckd-stage
get_ckd_stage <- function(egfr) {
  if (is.na(egfr)) {
    return(NA)
  } else if (egfr >= 90) {
    return(1)
  } else if (egfr >= 60) {
    return(2)
  } else if (egfr >= 30) {
    return(3)
  } else if (egfr >= 15) {
    return(4)
  } else {
    return(5)
  }
}
```

## 2.5 ckd_epi_gfr (CKD-EPI 2021 creatinine-only)

```{r}
#| label: fn-ckd-epi-gfr
ckd_epi_gfr <- function(creat, male, age) {
  k <- ifelse(male == 1, 0.9, 0.7)
  a <- ifelse(male == 1, -0.302, -0.241)
  m <- ifelse(male == 1, 1, 1.012)
  gfr <- 142 * min(c(creat / k, 1))^a * max(c(creat / k, 1))^-1.2 * 0.9938^age * m
  return(gfr)
}

gfr_calc <- Vectorize(ckd_epi_gfr)
```

---

# 3 Variable Dictionary

| Source (`treated group.xlsx`) | In Code | Derivation |
|---|---|---|
| `empi` | `EMPI` | Renamed uppercase (RPDR) |
| `ICI_Dose_Date` | `Index_date` | Renamed |
| `Date_of_Birth` | `Date_of_Birth` | Used to derive `Age` |
| `Gender_Legal_Sex` | `Gender_Legal_Sex` | Used to derive `male` |
| `Race` | `Race` | As-is |
| `Ethnic_Group` | `Ethnic_Group` | As-is |
| `Cancer_Diagnosis` | `Cancer_Diagnosis` | As-is |
| `Cancers_stage` | `Cancers_stage` | As-is |
| `ICI_Name` | `ICI_Name` | As-is |
| `Date_Of_Death` | `Date_Of_Death` | Used to derive `death`, `time_to_death` |
| `last_cre_date` | `last_cre_date` | Last creatinine lab date; used as follow-up date |

### Derived variables

| Variable | Description |
|---|---|
| `Age` | `floor((Index_date − Date_of_Birth) / 365.25)` |
| `male` | 1 if Male, else 0 |
| `last_cre_days` | Days from `Index_date` to `last_cre_date` |
| `death` | 1 if `Date_Of_Death` is not NA |
| `time_to_death` | `Date_Of_Death − Index_date` (days), or `last_cre_days` if alive |
| `death_Xyear` (1–10) | 1 if died within X×365 days |
| `survival_Xyear` (1–10) | 1 if `last_cre_days ≥ 365×X` |
| `pre_median_CRE_90days` | Median CRE ≤90 days pre-index (RPDR) |
| `eGFR_CRE_baseline` | CKD-EPI 2021 from baseline CRE |
| `ckd_stage_baseline` | CKD stages 1–5 |
| `ckd_stage_baseline_3` | 3-level: stage 1 / stage 2 / stage 3 or more |
| `pre_median_CRE_Xyear` (1–10) | Yearly median CRE (90-day window) |
| `eGFR_CRE_Xyear` (1–10) | Yearly eGFR |
| `ckd_stage_Xyear` (1–10) | Yearly CKD stage |
| `lab_num_Xyear` (1–10) | Number of CRE labs in yearly window |

> Follow-up uses `last_cre_date` → `last_cre_days` for censoring.

---

# 4 Data Loading

## 4.1 Load treated group Excel

```{r}
#| label: load-treated
treated <- read_excel("treated group.xlsx")
```

## 4.2 Rename & derive variables

```{r}
#| label: derive-vars
# last_cre_date: date of last creatinine lab (from source Excel)
# last_cre_days: derived days from Index_date to last_cre_date (used for censoring)
treated <- treated %>%
  rename(
    EMPI       = empi,
    Index_date = ICI_Dose_Date
  ) %>%
  mutate(
    Index_date    = as.Date(Index_date),
    Date_of_Birth = as.Date(Date_of_Birth),
    Date_Of_Death = as.Date(Date_Of_Death),
    last_cre_date = as.Date(last_cre_date),
    Age           = as.numeric(floor(difftime(Index_date, Date_of_Birth, units = "days") / 365.25)),
    male          = ifelse(Gender_Legal_Sex == "Male", 1, 0),
    last_cre_days = as.numeric(difftime(last_cre_date, Index_date, units = "days")),
    death         = ifelse(!is.na(Date_Of_Death), 1, 0),
    death_days    = as.numeric(difftime(Date_Of_Death, Index_date, units = "days")),
    time_to_death = death_days,
    time_to_death_status = ifelse(!is.na(death_days), 2, 0)
  ) %>%
  filter(last_cre_days > 0)
```

## 4.3 Death flags (1–10 years)

Uses `last_date` logic from Kavita master builder: `last_date = if_else(is.na(Date_Of_Death), last_cre_date, Date_Of_Death)`.

```{r}
#| label: death-flags
treated <- treated %>%
  mutate(
    last_date = if_else(is.na(Date_Of_Death), last_cre_date, Date_Of_Death),
    last_days = as.numeric(last_date - Index_date)
  )

for (yr in 1:10) {
  treated[[paste0("death_", yr, "year")]] <-
    ifelse(treated$last_days < 365 * yr, 1, 0)
}
```

## 4.4 Load RPDR files (Lab, Dia, Med)

```{r}
#| label: load-rpdr
rpdr_path <- "/Users/to909/Partners HealthCare Dropbox/Tianqi Ouyang/CKD after ICI/RPDR/"

lab <- read.table(
  paste0(rpdr_path, "mes98_093025104659807332_Lab.txt"),
  sep = "|", fill = TRUE, quote = "", header = TRUE
) %>%
  select(EMPI, EPIC_PMRN, Seq_Date_Time, Result, Group_Id, Test_Description)
lab$Seq_Date_Time <- as.Date(lab$Seq_Date_Time, "%m/%d/%Y")

dia <- read.table(
  paste0(rpdr_path, "mes98_093025104659807332_Dia.txt"),
  sep = "|", fill = TRUE, quote = "", header = TRUE
) %>%
  select(EMPI, EPIC_PMRN, Code, Date, Diagnosis_Name)
dia$Date <- as.Date(dia$Date, "%m/%d/%Y")

med <- read.table(
  paste0(rpdr_path, "mes98_093025104659807332_Med.txt"),
  sep = "|", fill = TRUE, quote = "", header = TRUE
) %>%
  select(EMPI, Medication_Date, Medication)
med$Medication_Date <- as.Date(med$Medication_Date, "%m/%d/%Y")
```

## 4.5 Create master_lab / master_dia / master_med

```{r}
#| label: create-master-lab
master_list <- treated %>% select(EMPI, Index_date)
master_lab  <- lab %>% merge(master_list, by = "EMPI")
master_dia  <- dia %>% merge(master_list, by = "EMPI")
master_med  <- med %>% merge(master_list, by = "EMPI")
```

## 4.6 RPDR-derived functions

```{r}
#| label: fn-rpdr-helpers
med_list <- function(names, index_date, excluded_names = NULL, data) {
  if (!is.null(excluded_names)) {
    result <- data %>%
      filter(Medication_Date <= {{index_date}}) %>%
      filter(
        grepl(paste(names, collapse = "|"), Medication, ignore.case = TRUE),
        !grepl(paste(excluded_names, collapse = "|"), Medication, ignore.case = TRUE)
      ) %>%
      distinct(EMPI)
  } else {
    result <- data %>%
      filter(Medication_Date <= {{index_date}}) %>%
      filter(grepl(paste(names, collapse = "|"), Medication, ignore.case = TRUE)) %>%
      distinct(EMPI)
  }
  return(result)
}

dia_list_icd <- function(data, index_date, icd_code, excluded_names = NULL,
                         med_data = NULL, med_names = NULL) {
  if (is.null(excluded_names)) {
    dia_result <- data %>%
      filter(Date <= {{index_date}}) %>%
      filter(grepl(paste(icd_code, collapse = "|"), Code, ignore.case = TRUE)) %>%
      distinct(EMPI)
  } else {
    dia_result <- data %>%
      filter(Date <= {{index_date}}) %>%
      filter(
        grepl(paste(icd_code, collapse = "|"), Code, ignore.case = TRUE),
        !grepl(paste(excluded_names, collapse = "|"), Diagnosis_Name, ignore.case = TRUE)
      ) %>%
      distinct(EMPI)
  }
  if (!is.null(med_data)) {
    med_result <- med_data %>%
      filter(Medication_Date <= {{index_date}}) %>%
      filter(grepl(paste(med_names, collapse = "|"), Medication, ignore.case = TRUE)) %>%
      distinct(EMPI)
  } else {
    med_result <- NULL
  }
  rbind(dia_result, med_result) %>% distinct(EMPI)
}

dia_list_name <- function(data, listed_names, index_date, excluded_names = NULL,
                          med_data = NULL, med_names = NULL) {
  if (is.null(excluded_names)) {
    dia_result <- data %>%
      filter(Date <= {{index_date}}) %>%
      filter(grepl(paste(listed_names, collapse = "|"), Diagnosis_Name, ignore.case = TRUE)) %>%
      distinct(EMPI)
  } else {
    excluded_pattern <- paste(gsub("([][()])", "\\\\\\1", excluded_names), collapse = "|")
    dia_result <- data %>%
      filter(Date <= {{index_date}}) %>%
      filter(
        grepl(paste(listed_names, collapse = "|"), Diagnosis_Name, ignore.case = TRUE),
        !grepl(excluded_pattern, Diagnosis_Name, ignore.case = TRUE)
      ) %>%
      distinct(EMPI)
  }
  if (!is.null(med_data)) {
    med_result <- med_data %>%
      filter(Medication_Date <= {{index_date}}) %>%
      filter(grepl(paste(med_names, collapse = "|"), Medication, ignore.case = TRUE)) %>%
      distinct(EMPI)
  } else {
    med_result <- NULL
  }
  rbind(dia_result, med_result) %>% distinct(EMPI)
}

dia_list_icd_after_index <- function(data, index_date, icd_code, excluded_names = NULL) {
  if (is.null(excluded_names)) {
    result <- data %>%
      filter(Date > {{index_date}}) %>%
      filter(grepl(paste(icd_code, collapse = "|"), Code, ignore.case = TRUE))
  } else {
    result <- data %>%
      filter(Date > {{index_date}}) %>%
      filter(
        grepl(paste(icd_code, collapse = "|"), Code, ignore.case = TRUE),
        !grepl(paste(excluded_names, collapse = "|"), Diagnosis_Name, ignore.case = TRUE)
      )
  }
  result %>%
    arrange(EMPI, Date) %>%
    distinct(EMPI, .keep_all = TRUE) %>%
    select(EMPI, first_event_date = Date)
}

make_ckd_outcome_date <- function(data, id_var, date_var, condition_expr,
                                   days_threshold = 90, outcome_name = "default") {
  id_var      <- enquo(id_var)
  date_var    <- enquo(date_var)
  cond_q      <- enquo(condition_expr)
  outcome_sym <- ensym(outcome_name)
  outcome_date_name <- paste0(outcome_name, "_date")

  working <- data %>%
    arrange(!!id_var, !!date_var) %>%
    group_by(!!id_var) %>%
    mutate(
      .cond   = if_else(!!cond_q, 1L, 0L),
      .change = .cond != lag(.cond, default = first(.cond)),
      .run    = cumsum(.change)
    )

  episodes <- working %>%
    group_by(!!id_var, .run) %>%
    summarise(
      .code = first(.cond), .start = first(!!date_var),
      .last = last(!!date_var), .groups = "drop"
    ) %>%
    group_by(!!id_var) %>%
    mutate(.end = lead(.start),
           .end = if_else(is.na(.end), .last, .end),
           .days = as.integer(.end - .start)) %>%
    ungroup()

  episodes %>%
    filter(!is.na(.days)) %>%
    group_by(!!id_var) %>%
    summarise(
      has_event  = any(.code == 1L & .days >= days_threshold, na.rm = TRUE),
      event_date = if (any(.code == 1L & .days >= days_threshold, na.rm = TRUE)) {
        min(.start[.code == 1L & .days >= days_threshold], na.rm = TRUE)
      } else { as.Date(NA) },
      .groups = "drop"
    ) %>%
    mutate(!!outcome_sym := as.integer(has_event)) %>%
    rename(!!outcome_date_name := event_date) %>%
    select(-has_event)
}
```

## 4.7 Derive comorbidities & medications from RPDR

```{r}
#| label: derive-comorbidities
# Medication name lists
ppi_names <- c("omeprazole", "Prilosec", "lansoprazole", "Prevacid",
               "dexlansoprazole", "rabeprazole", "Aciphex",
               "pantoprazole", "Protonix", "esomeprazole", "Nexium")
ace_arb_names <- c("benazepril", "captopril", "enalapril", "fosinopril",
                   "lisinopril", "ramipril", "losartan", "valsartan",
                   "irbesartan", "olmesartan", "candesartan", "telmisartan",
                   "azilsartan", "eprosartan")
diuretics_names <- c("bumetanide", "furosemide", "torsemide",
                     "hydrochlorothiazide", "chlorthalidone", "metolazone",
                     "spironolactone", "triamterene", "amiloride", "eplerenone")
steroids_names <- c("prednisone", "methylprednisolone", "prednisolone",
                    "dexamethasone", "budesonide")
no_steroid_names <- c("Inhaler", "Nebulization", "Ointment", "Ophthalmic")
statins_names <- c("Atorvastatin", "Lipitor", "Fluvastatin", "Lovastatin",
                   "Pitavastatin", "Pravastatin", "Rosuvastatin", "Simvastatin")

# Diagnosis codes
diabetes_codes <- c("250.00", "250.01", "250.02", "250.03", "E10.9", "E11.9",
                    "E10.65", "E11.65", "E10.21", "E11.21", "E10.22", "E11.22",
                    "E13.9")
htn_codes <- c("401.1", "401.0", "401.9", "I10", "I15.1", "I15.9")
no_htn_list <- c("Pulmonary Embolism", "Borderline hypertension",
                 "Chronic venous hypertension")
htn_med_list <- c("hydrochlorothiazide", "amlodipine", "lisinopril",
                  "losartan", "valsartan", "diltiazem", "clonidine")
cad_list <- c("Coronary artery disease", "Coronary atherosclerosis",
              "Atherosclerotic heart disease", "Angina pectoris",
              "Unstable angina")
smoking_list <- c("smoking", "tobacco", "cigarette", "nicotine", "smoke")
esrd_kt_codes <- c("585.6", "N18.6", "V45.1", "Z94.0", "Z99.2",
                   "996.81", "T86.1", "T86.10", "T86.11", "T86.12")

# Derive
ppi      <- med_list(data = master_med, names = ppi_names, index_date = Index_date)
ace_arb  <- med_list(data = master_med, names = ace_arb_names, index_date = Index_date)
diu      <- med_list(data = master_med, names = diuretics_names, index_date = Index_date)
steroids <- med_list(data = master_med, names = steroids_names,
                     index_date = Index_date, excluded_names = no_steroid_names)
statins  <- med_list(data = master_med, names = statins_names, index_date = Index_date)
esrd_kt  <- dia_list_icd(master_dia, icd_code = esrd_kt_codes, index_date = Index_date)
dm  <- dia_list_icd(master_dia, icd_code = diabetes_codes, index_date = Index_date)
htn <- dia_list_icd(master_dia, icd_code = htn_codes, excluded_names = no_htn_list,
                    med_data = master_med, med_names = htn_med_list,
                    index_date = Index_date)
cad <- dia_list_name(master_dia, listed_names = cad_list, index_date = Index_date)
cirrhosis <- master_dia %>%
  filter(Date <= Index_date) %>%
  filter(grepl("cirrhosis", Diagnosis_Name, ignore.case = TRUE)) %>%
  distinct(EMPI)
smoking_dx <- master_dia %>%
  filter(grepl(paste(smoking_list, collapse = "|"), Diagnosis_Name, ignore.case = TRUE)) %>%
  distinct(EMPI)

treated <- treated %>%
  mutate(
    diu       = ifelse(EMPI %in% diu$EMPI, 1, 0),
    ace_arb   = ifelse(EMPI %in% ace_arb$EMPI, 1, 0),
    esrd_kt   = ifelse(EMPI %in% esrd_kt$EMPI, 1, 0),
    dm        = ifelse(EMPI %in% dm$EMPI, 1, 0),
    htn       = ifelse(EMPI %in% htn$EMPI, 1, 0),
    ppi       = ifelse(EMPI %in% ppi$EMPI, 1, 0),
    steroids  = ifelse(EMPI %in% steroids$EMPI, 1, 0),
    smoking   = ifelse(EMPI %in% smoking_dx$EMPI, 1, 0),
    cad       = ifelse(EMPI %in% cad$EMPI, 1, 0),
    cirrhosis = ifelse(EMPI %in% cirrhosis$EMPI, 1, 0),
    statins   = ifelse(EMPI %in% statins$EMPI, 1, 0)
  )
```

## 4.8 ICI_Class

```{r}
#| label: derive-ici-class
treated <- treated %>%
  mutate(
    ICI_Class = case_when(
      ICI_Name %in% c("pembrolizumab", "nivolumab", "cemiplimab",
                       "Nivolumab", "Dostarlimab", "Cemiplimab") ~ "PD1",
      ICI_Name %in% c("atezolizumab", "avelumab", "durvalumab") ~ "PDL1",
      ICI_Name == "ipilimumab" ~ "CTLA4",
      ICI_Name == "ipilimumab AND nivolumab" ~ "Combination",
      TRUE ~ NA_character_
    )
  )
```

---

# 5 Baseline Lab Values

## 5.1 Baseline creatinine (median within 90 days pre-index)

```{r}
#| label: baseline-cre
pre_lab_baseline <- pre_lab_median(
  master_lab, month = FALSE, time = 90,
  index_date = Index_date, Group_Id = "CRE"
)

treated <- treated %>%
  merge(pre_lab_baseline, by = "EMPI", all.x = TRUE)
```

## 5.2 Baseline eGFR & CKD stage

```{r}
#| label: baseline-egfr
treated$eGFR_CRE_baseline <- mapply(
  gfr_calc,
  treated$pre_median_CRE_90days,
  treated$male,
  treated$Age
)

treated <- treated %>%
  mutate(
    ckd_stage_baseline = case_when(
      eGFR_CRE_baseline >= 90                           ~ 1,
      eGFR_CRE_baseline >= 60 & eGFR_CRE_baseline < 90 ~ 2,
      eGFR_CRE_baseline >= 30 & eGFR_CRE_baseline < 60 ~ 3,
      eGFR_CRE_baseline >= 15 & eGFR_CRE_baseline < 30 ~ 4,
      eGFR_CRE_baseline < 15                            ~ 5
    ),
    ckd_stage_baseline_3 = case_when(
      ckd_stage_baseline == 1  ~ "stage 1",
      ckd_stage_baseline == 2  ~ "stage 2",
      ckd_stage_baseline >= 3  ~ "stage 3 or more"
    )
  )
```

## 5.3 CKD outcomes from RPDR Lab (sustained eGFR-based)

```{r}
#| label: clean-ckd-outcomes
outcome_labs <- master_lab %>%
  select(EMPI, Seq_Date_Time, Result, Group_Id, Index_date) %>%
  merge(treated %>% select(EMPI, pre_median_CRE_90days, eGFR_CRE_baseline,
                            male, Age),
        by = "EMPI") %>%
  filter(Group_Id == "CRE", grepl("^\\d+\\.\\d+$", Result)) %>%
  filter(!is.na(Result)) %>%
  mutate(Result = as.numeric(Result)) %>%
  filter(Seq_Date_Time > Index_date) %>%
  mutate(eGFR_CRE = mapply(gfr_calc, Result, male, Age)) %>%
  select(EMPI, eGFR_CRE_baseline, eGFR_CRE, eGFR_CRE_date = Seq_Date_Time) %>%
  mutate(pct_drop = round((eGFR_CRE_baseline - eGFR_CRE) / eGFR_CRE_baseline, 3)) %>%
  arrange(EMPI, eGFR_CRE_date)

ckd_incidence <- make_ckd_outcome_date(
  outcome_labs, EMPI, eGFR_CRE_date,
  pct_drop > 0.25 & eGFR_CRE < 60 & eGFR_CRE_baseline >= 60,
  90, "ckd_incidence")

eskd_defination_1 <- make_ckd_outcome_date(
  outcome_labs, EMPI, eGFR_CRE_date,
  eGFR_CRE < 10, 90, "eskd_defination_1")

ckd_progression <- make_ckd_outcome_date(
  outcome_labs, EMPI, eGFR_CRE_date,
  pct_drop > 0.30 & eGFR_CRE_baseline < 60, 90, "ckd_progression")

treated <- treated %>%
  merge(ckd_incidence, by = "EMPI", all.x = TRUE) %>%
  merge(eskd_defination_1, by = "EMPI", all.x = TRUE) %>%
  merge(ckd_progression, by = "EMPI", all.x = TRUE) %>%
  mutate(
    ckd_incidence     = replace_na(ckd_incidence, 0),
    eskd_defination_1 = replace_na(eskd_defination_1, 0),
    ckd_progression   = replace_na(ckd_progression, 0),
    esrd_kt_after_ici = ifelse(EMPI %in% (dia_list_icd_after_index(
      master_dia, icd_code = esrd_kt_codes, index_date = Index_date)$EMPI), 1, 0),
    eskd_composite    = ifelse(esrd_kt_after_ici == 1 | eskd_defination_1 == 1, 1, 0),
    ckd_composite     = ifelse(eskd_composite == 1 | ckd_incidence == 1 | ckd_progression == 1, 1, 0),
    time_to_ckd_incidence   = as.numeric(ckd_incidence_date - Index_date),
    time_to_ckd_progression = as.numeric(ckd_progression_date - Index_date),
    time_to_eskd_composite  = as.numeric(pmin(eskd_defination_1_date, na.rm = TRUE) - Index_date),
    time_to_ckd_composite   = pmin(time_to_ckd_incidence, time_to_ckd_progression,
                                    time_to_eskd_composite, na.rm = TRUE)
  )
```

---

# 6 Yearly Lab Values & Survival Flags

## 6.1 Survival flags (1–10 years)

```{r}
#| label: survival-flags
for (yr in 1:10) {
  treated[[paste0("survival_", yr, "year")]] <-
    ifelse(treated$last_cre_days >= 365 * yr, 1, 0)
}
```

## 6.2 Yearly CRE medians, eGFR, CKD stage (1–10 years)

```{r}
#| label: yearly-cre
yearly_cre_list <- list()
for (yr in 1:10) {
  yearly_cre_list[[yr]] <- pre_lab_median_year(
    master_lab, year = yr, days = 90,
    index_date = Index_date, Group_Id = "CRE"
  )
}

treated <- left_join_all(treated, yearly_cre_list, by = "EMPI")

for (yr in 1:10) {
  cre_col   <- paste0("pre_median_CRE_", yr, "year")
  egfr_col  <- paste0("eGFR_CRE_", yr, "year")
  stage_col <- paste0("ckd_stage_", yr, "year")

  treated[[egfr_col]]  <- mapply(gfr_calc, treated[[cre_col]],
                                  treated$male, treated$Age)
  treated[[stage_col]] <- sapply(treated[[egfr_col]], get_ckd_stage)
}
```

---

## 6.3 AKI (from RPDR Lab)

```{r}
#| label: clean-aki
aki_data <- master_lab %>%
  merge(pre_lab_baseline, by = "EMPI", all.x = TRUE) %>%
  filter(Group_Id == "CRE", grepl("^\\d+\\.\\d+$", Result)) %>%
  filter(!is.na(Result), !is.na(pre_median_CRE_90days)) %>%
  mutate(Result = as.numeric(Result)) %>%
  filter(Seq_Date_Time > Index_date,
         Seq_Date_Time <= Index_date + 90) %>%
  filter(Result > 1.5 * pre_median_CRE_90days) %>%
  arrange(EMPI, Seq_Date_Time) %>%
  distinct(EMPI, .keep_all = TRUE) %>%
  select(EMPI, aki_date = Seq_Date_Time)

treated <- treated %>%
  merge(aki_data, by = "EMPI", all.x = TRUE) %>%
  mutate(aki = ifelse(!is.na(aki_date), 1, 0))
```

## 6.4 Competing risk & analysis variables

```{r}
#| label: competing-risk-vars
treated <- treated %>%
  mutate(
    year_10 = Age / 10,
    eGFR_CRE_baseline_decrease_10 = (eGFR_CRE_baseline * -1) / 10,
    # Competing risk: pmin of all event times, then priority-based status
    time_to_ckd_cmp = pmin(time_to_ckd_composite, last_cre_days, death_days, na.rm = TRUE),
    status_ckd_cmp = case_when(
      # Priority 1: Death (2)
      !is.na(death_days) & time_to_ckd_cmp == death_days ~ 2,
      # Priority 2: CKD Event (1)
      !is.na(time_to_ckd_composite) & time_to_ckd_cmp == time_to_ckd_composite ~ 1,
      # Priority 3: Censored / Last Creatinine (0)
      !is.na(last_cre_days) & time_to_ckd_cmp == last_cre_days ~ 0,
      TRUE ~ NA_real_
    ),
    status_ckd_cmp = factor(status_ckd_cmp, levels = c(0, 1, 2),
                             labels = c("Censored", "CKD Composite", "Death")),
    aki_before_ckd = ifelse(as.numeric(aki_date - Index_date) <= time_to_ckd_composite, 1, 0)
  )

# Yearly outcome flags
for (yr in 1:10) {
  treated[[paste0("ckd_incidence_", yr, "year")]] <-
    ifelse(treated$time_to_ckd_incidence < 365 * yr, 1, 0)
  treated[[paste0("ckd_progression_", yr, "year")]] <-
    ifelse(treated$time_to_ckd_progression < 365 * yr, 1, 0)
  treated[[paste0("eskd_composite_", yr, "year")]] <-
    ifelse(treated$time_to_eskd_composite < 365 * yr, 1, 0)
  treated[[paste0("ckd_composite_", yr, "year")]] <-
    ifelse(treated$time_to_ckd_composite < 365 * yr, 1, 0)
}

treated <- treated %>%
  mutate(across(
    .cols = matches("(ckd_incidence|ckd_progression|eskd_composite|ckd_composite).*year$"),
    .fns  = ~ factor(replace_na(., 0))
  ))
```

---

# 7 Exclusion & Final Dataset

## 7.1 Exclude missing baseline CRE and eGFR < 10

```{r}
#| label: exclusion
n_before <- nrow(treated)

treated <- treated %>%
  filter(!is.na(pre_median_CRE_90days)) %>%
  filter(eGFR_CRE_baseline >= 10)

n_after <- nrow(treated)
cat("Excluded", n_before - n_after, "patients.",
    n_after, "patients remain.\n")
```

## 7.2 Save treated_group.xlsx

```{r}
#| label: save-xlsx
write_xlsx(treated, "treated_group.xlsx")
cat("Saved treated_group.xlsx with", nrow(treated), "rows and",
    ncol(treated), "columns.\n")
```

---

:::

# Descriptive Analysis {.unnumbered}

## Table 1 – Demographics & Cancer {.unnumbered}

```{r}
#| label: table1-demographics
#| include: true
treated %>%
  select(
    Age, Gender_Legal_Sex, Race, Ethnic_Group,
    ICI_Name, ICI_Class,
    ## Comorbidities
    dm, htn, cad, cirrhosis,
    ## Medications
    ace_arb, diu, ppi, steroids, smoking, statins,
    ## Kidney
    eGFR_CRE_baseline, ckd_stage_baseline,
    ## AKI
    aki
  ) %>%
  tbl_summary(
    type = list(Age ~ "continuous", eGFR_CRE_baseline ~ "continuous"),
    statistic = list(all_continuous() ~ "{median} ({p25}, {p75})"),
    missing_text = "Missing"
  ) %>%
  bold_labels()
```

## Baseline Kidney Function {.unnumbered}

### eGFR distribution

```{r}
#| label: baseline-egfr-dist
#| include: true
#| fig-cap: "Distribution of baseline eGFR (CKD-EPI 2021)"
ggplot(treated, aes(x = eGFR_CRE_baseline)) +
  geom_histogram(binwidth = 10, fill = "steelblue", color = "white") +
  labs(x = "Baseline eGFR (mL/min/1.73m²)", y = "Count",
       title = "Distribution of Baseline eGFR") +
  theme_minimal()
```

### CKD stage distribution

```{r}
#| label: baseline-ckd-stage
#| include: true
treated %>%
  select(ckd_stage_baseline, ckd_stage_baseline_3) %>%
  tbl_summary(
    label = list(
      ckd_stage_baseline   ~ "CKD Stage (1-5)",
      ckd_stage_baseline_3 ~ "CKD Stage (3-level)"
    ),
    missing = "ifany"
  ) %>%
  bold_labels()
```

## Yearly eGFR Trajectory {.unnumbered}

```{r}
#| label: egfr-trajectory
#| include: true
#| fig-cap: "Median eGFR by year post-ICI (with IQR)"
egfr_cols <- paste0("eGFR_CRE_", 1:10, "year")

egfr_summary <- tibble(
  year = 1:10,
  median_egfr = sapply(egfr_cols, function(col) median(treated[[col]], na.rm = TRUE)),
  q25         = sapply(egfr_cols, function(col) quantile(treated[[col]], 0.25, na.rm = TRUE)),
  q75         = sapply(egfr_cols, function(col) quantile(treated[[col]], 0.75, na.rm = TRUE)),
  n           = sapply(egfr_cols, function(col) sum(!is.na(treated[[col]])))
)

ggplot(egfr_summary, aes(x = year, y = median_egfr)) +
  geom_ribbon(aes(ymin = q25, ymax = q75), alpha = 0.2, fill = "steelblue") +
  geom_line(color = "steelblue", linewidth = 1) +
  geom_point(color = "steelblue", size = 2) +
  geom_text(aes(label = paste0("n=", n)), vjust = -1.5, size = 3) +
  scale_x_continuous(breaks = 1:10) +
  labs(x = "Year Post-ICI", y = "eGFR (mL/min/1.73m²)",
       title = "Median eGFR Trajectory (with IQR)") +
  theme_minimal()
```

## Yearly CKD Stage Distribution {.unnumbered}

```{r}
#| label: yearly-ckd-stage
#| include: true
#| fig-cap: "CKD stage distribution by year post-ICI"
stage_data <- map_dfr(1:10, function(yr) {
  col <- paste0("ckd_stage_", yr, "year")
  tibble(
    year      = yr,
    ckd_stage = factor(treated[[col]])
  ) %>%
    filter(!is.na(ckd_stage))
})

ggplot(stage_data, aes(x = factor(year), fill = ckd_stage)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(palette = "RdYlGn", direction = -1, name = "CKD Stage") +
  labs(x = "Year Post-ICI", y = "Proportion",
       title = "CKD Stage Distribution by Year") +
  theme_minimal()
```

## Overall Survival (Kaplan-Meier) {.unnumbered}

```{r}
#| label: km-survival
#| include: true
#| fig-cap: "Kaplan-Meier curve – overall survival from ICI start"
surv_obj <- Surv(time = treated$last_days, event = treated$death)
km_fit   <- survfit(surv_obj ~ 1)

ggsurvplot(
  km_fit,
  data        = treated,
  risk.table  = TRUE,
  xlab        = "Days from ICI Start",
  ylab        = "Survival Probability",
  title       = "Overall Survival – Treated Group",
  palette     = "steelblue",
  ggtheme     = theme_minimal()
)
```

## Death by Year {.unnumbered}

```{r}
#| label: death-flags-table
#| include: true
death_counts <- tibble(
  Year  = 1:10,
  Deaths = sapply(1:10, function(yr) {
    sum(treated[[paste0("death_", yr, "year")]] == 1, na.rm = TRUE)
  }),
  Total  = nrow(treated),
  Pct    = round(Deaths / Total * 100, 1)
)

knitr::kable(death_counts, col.names = c("Year", "Cumulative Deaths", "Total N", "% Died"),
             caption = "Cumulative deaths by year post-ICI")
```

## Survival by Year {.unnumbered}

```{r}
#| label: survival-flags-table
#| include: true
survival_counts <- tibble(
  Year     = 1:10,
  Survived = sapply(1:10, function(yr) {
    sum(treated[[paste0("survival_", yr, "year")]] == 1, na.rm = TRUE)
  }),
  Total    = nrow(treated),
  Pct      = round(Survived / Total * 100, 1)
)

knitr::kable(survival_counts,
             col.names = c("Year", "Survived to Year", "Total N", "% Survived"),
             caption = "Patients surviving to each yearly landmark")
```

## CKD Outcomes Summary {.unnumbered}

```{r}
#| label: outcome-summary
#| include: true
outcome_counts <- tibble(
  Year = 1:10,
  `CKD Incidence` = sapply(1:10, function(yr) {
    col <- paste0("ckd_incidence_", yr, "year")
    sum(treated[[col]] == 1, na.rm = TRUE)
  }),
  `CKD Progression` = sapply(1:10, function(yr) {
    col <- paste0("ckd_progression_", yr, "year")
    sum(treated[[col]] == 1, na.rm = TRUE)
  }),
  `ESKD Composite` = sapply(1:10, function(yr) {
    col <- paste0("eskd_composite_", yr, "year")
    sum(treated[[col]] == 1, na.rm = TRUE)
  }),
  `CKD Composite` = sapply(1:10, function(yr) {
    col <- paste0("ckd_composite_", yr, "year")
    sum(treated[[col]] == 1, na.rm = TRUE)
  }),
  `N Followed` = sapply(1:10, function(yr) {
    sum(treated[[paste0("survival_", yr, "year")]] == 1, na.rm = TRUE)
  })
)

knitr::kable(outcome_counts,
             caption = "Cumulative CKD outcomes by year post-ICI")
```

## Cumulative Incidence – CKD Composite vs Death {.unnumbered}

```{r}
#| label: cif-plot
#| include: true
#| fig-cap: "Cumulative incidence of CKD composite and death (competing risks)"
cuminc(Surv(time_to_ckd_cmp, status_ckd_cmp) ~ 1, data = treated) %>%
  ggcuminc(outcome = c("CKD Composite", "Death")) +
  add_confidence_interval() +
  add_risktable() +
  labs(x = "Years", y = "Cumulative Incidence") +
  scale_ggsurvfit() +
  scale_x_continuous(
    breaks = seq(0, 365 * 10, by = 365),
    labels = 0:10,
    limits = c(0, 365 * 10)
  )
```

## Univariate Fine-Gray Competing Risk Model {.unnumbered}

```{r}
#| label: fg-univariate
#| include: true
# Select candidate predictors, drop any with <2 unique non-NA values
fg_vars <- c("year_10", "male", "Race", "dm", "htn", "cad", "cirrhosis",
             "ace_arb", "diu", "ppi", "steroids", "smoking", "statins",
             "eGFR_CRE_baseline_decrease_10", "ICI_Class")
# Keep only variables with >=2 unique non-NA values
fg_vars_keep <- fg_vars[sapply(fg_vars, function(v) {
  length(unique(na.omit(treated[[v]]))) >= 2
})]
fg_data <- treated %>%
  select(all_of(fg_vars_keep), time_to_ckd_cmp, status_ckd_cmp)

fg_data %>%
  tbl_uvregression(
    method = tidycmprsk::crr,
    method.args = list(failcode = "CKD Composite", cencode = "Censored"),
    y = Surv(time_to_ckd_cmp, status_ckd_cmp),
    pvalue_fun = ~style_pvalue(.x, digits = 3),
    exponentiate = TRUE
  ) %>%
  bold_p(0.05) %>%
  bold_labels() %>%
  add_n()
```

## Multivariate Fine-Gray Model {.unnumbered}

```{r}
#| label: fg-multivariate
#| include: true
# Build formula dynamically — drop ICI_Class if <2 levels
mv_vars <- c("year_10", "male", "dm", "htn", "cad", "ppi", "smoking",
             "eGFR_CRE_baseline_decrease_10", "cirrhosis", "statins")
if (is.factor(treated$ICI_Class) && nlevels(droplevels(treated$ICI_Class)) >= 2) {
  mv_vars <- c(mv_vars, "ICI_Class")
}
mv_formula <- as.formula(paste("Surv(time_to_ckd_cmp, status_ckd_cmp) ~",
                                paste(mv_vars, collapse = " + ")))

tidycmprsk::crr(
  mv_formula,
  failcode = "CKD Composite", cencode = "Censored",
  data = treated
) %>%
  tbl_regression(exp = TRUE) %>%
  add_n() %>%
  bold_p(0.05) %>%
  bold_labels()
```

## Univariate Cause-Specific Hazard Model {.unnumbered}

```{r}
#| label: csh-univariate
#| include: true
treated <- treated %>%
  mutate(status_ckd_csh = ifelse(status_ckd_cmp == "CKD Composite", 1, 0))

treated %>%
  select(all_of(fg_vars_keep), time_to_ckd_cmp, status_ckd_csh) %>%
  tbl_uvregression(
    method = survival::coxph,
    y = Surv(time_to_ckd_cmp, status_ckd_csh),
    exponentiate = TRUE
  ) %>%
  bold_p(0.05) %>%
  bold_labels() %>%
  add_n()
```

## CIF by Diabetes Status {.unnumbered}

```{r}
#| label: cif-dm
#| include: true
#| fig-cap: "Cumulative incidence of CKD composite by diabetes status"
treated <- treated %>%
  mutate(
    time_to_ckd_cmp_year = time_to_ckd_cmp / 365,
    dm_label = factor(dm, levels = c(0, 1), labels = c("No Diabetes", "Diabetes"))
  )

cuminc(Surv(time_to_ckd_cmp_year, status_ckd_cmp) ~ dm_label, data = treated) %>%
  ggcuminc(outcome = "CKD Composite") +
  add_confidence_interval() +
  add_risktable(
    times = seq(0, 10, by = 1),
    risktable_stats = c("n.risk", "cum.event"),
    stats_label = list(n.risk = "At Risk", cum.event = "Events")
  ) +
  scale_color_manual(values = c("Diabetes" = "#619CFF", "No Diabetes" = "#F8766D")) +
  scale_fill_manual(values = c("Diabetes" = "#619CFF", "No Diabetes" = "#F8766D")) +
  scale_x_continuous(breaks = seq(0, 10, by = 1), limits = c(0, 10)) +
  labs(x = "Years since ICI", y = "Cumulative Incidence",
       title = "CIF of CKD Composite by DM Status") +
  theme_classic() +
  theme(legend.position = "bottom")
```
